---
title: Improving forecasts via subspace projections
author: Rob J Hyndman
date: 19 September 2024
toc: true
format:
  presentation-beamer:
    pdf-engine: pdflatex
    knitr:
      opts_chunk:
        dev: "CairoPDF"
    template-partials:
      - before-title.tex
    include-in-header: header.tex
    keep-tex: true
cite-method: biblatex
biblatexoptions: natbib,style=authoryear
bibliography: hts.bib
highlight-style: tango
execute:
  echo: false
  message: false
  warning: false
  cache: true
abstract: Univariate, multivariate, and hierarchical forecasts can all be improved using projections onto linear subspaces, regardless of what forecasting method is used. I will show some theoretical guarantees of this statement, and demonstrate using empirical applications how linear projections can lead to (sometimes dramatic) improvements in forecast accuracy.
---

```{r}
source("setup.R")
```

## Outline

\vspace*{0.7cm}\tableofcontents

# Improving hierarchical forecasts

## Australian tourism regions

```{r}
#| label: ausmap
#| fig-height: 3.5
library(sf)
# Use Okabe-Ito color-blind friendly color palette
state_colors <- c(
  `New South Wales` = "#56b4e9",
  `Victoria` = "#0072b2",
  `Queensland` = "#009e73",
  `South Australia` = "#f0e442",
  `Northern Territory` = "#d55e00",
  `Western Australia` = "#e69f00",
  `Tasmania` = "#cc79a7",
  `Australian Capital Territory` = "#cccccc"
)
read_sf("tourism/Tourism_Regions_2020.shp") |>
  rename(State = "STE_NAME16") |>
  ggplot() +
  geom_sf(aes(fill = State), alpha = 0.8) +
  theme_void() +
  theme(text = ggplot2::element_text(family = 'Fira Sans')) +
  scale_fill_manual(values = state_colors)
```

\only<2>{\begin{textblock}{6.5}(9.2,1.4)
\begin{block}{}%\fontsize{12}{13}\sf
  \begin{itemize}\itemsep=0cm\parskip=0cm
    \item Monthly data on visitor night from 1998 -- 2017
    \item From \textit{National Visitor Survey}, annual interviews of 120,000 Australians aged 15+.
    \item Geographical hierarchy split by
    \begin{itemize}
    \item 7 states
    \item 27 zones
    \item 75 regions
    \end{itemize}
  \end{itemize}
\end{block}
\end{textblock}}


## Australian tourism data

```{r}
#| label: tourism_plots
#| fig-width: 12
#| fig-height: 5
p1 <- tourism |>
  summarise(visitors = sum(visitors)) |>
  autoplot(visitors) +
  ylab("Overnight trips") + xlab("Month") +
  scale_y_log10() +
  ggtitle("Total domestic travel: Australia")
p2 <- tourism |>
  group_by(state) |>
  summarise(visitors = sum(visitors)) |>
  autoplot(visitors) +
  ylab("Overnight trips") + xlab("Month") +
  scale_y_log10() +
  ggtitle("Total domestic travel: by state") +
  scale_color_manual(
    values =
      c(
        NSW = "#56b4e9",
        VIC = "#0072b2",
        QLD = "#009e73",
        SA = "#f0e442",
        NT = "#d55e00",
        WA = "#e69f00",
        TAS = "#cc79a7",
        ACT = "#cccccc"
      )
  )
p3 <- tourism |>
  filter(state == "NSW") |>
  group_by(zone) |>
  summarise(visitors = sum(visitors)) |>
  mutate(zone = paste0("NSW/", zone)) |>
  autoplot(visitors) +
  ylab("Overnight trips") + xlab("Month") +
  scale_y_log10() +
  ggtitle("Total domestic travel: NSW by zone") +
  guides(colour = guide_legend(title = "state/zone"))
p4 <- tourism |>
  filter(zone == "South NSW") |>
  autoplot(visitors) +
  ylab("Overnight trips") + xlab("Month") +
  scale_y_log10() +
  ggtitle("Total domestic travel: South NSW by region")
aligned_plots <- align_patches(p1, p2, p3, p4)
for (i in seq_along(aligned_plots)) {
  fname <- paste0("./figs/tourism", i, ".pdf")
  cairo_pdf(fname, width = 12 * .8, height = 5 * .8)
  print(aligned_plots[[i]])
  crop::dev.off.crop(fname)
}
```

\only<1>{\placefig{0.1}{1.4}{width=15.8cm}{tourism1}}
\only<2>{\placefig{0.1}{1.4}{width=15.8cm}{tourism2}}
\only<3>{\placefig{0.1}{1.4}{width=15.8cm}{tourism3}}
\only<4>{\placefig{0.1}{1.4}{width=15.8cm}{tourism4}}

## Spectacle sales

\placefig{7}{4.9}{width=8.7cm}{spectacles}
\begin{textblock}{12}(0.3,1.2)
\begin{itemize}\tightlist
    \item Monthly UK sales data from 2000 -- 2014
    \item Provided by a large spectacle \rlap{manufacturer}
    \item Split by brand (26), gender (3), price range (6), materials (4), and stores (600)
    \item About 1 million bottom-level series
\end{itemize}
\end{textblock}

## Hierarchical time series
\vspace*{-0.2cm}

A \alert{\textbf{hierarchical time series}} is a collection of several time series that are linked together in a hierarchical structure.

\vspace*{-0.5cm}
\begin{center}
\begin{minipage}{9.6cm}
\begin{block}{}
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15]
\tikzstyle[level distance=.1cm]
\tikzstyle[sibling distance=7cm]
\tikzstyle{level 1}=[sibling distance=33mm,set style={{every node}+=[fill=blue!15]}]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AA}}
   child {node {AB}}
   child {node {AC}}
 }
 child {node {B}
   child {node {BA}}
   child {node {BB}}
   child {node {BC}}
 }
 child {node {C}
   child {node {CA}}
   child {node {CB}}
   child {node {CC}}
 };
\end{tikzpicture}
\end{block}
\end{minipage}
\end{center}

\pause\alert{Examples}\vspace*{-0.4cm}

 * Tourism by state and region
 * Retail sales by product groups, sub groups, and SKUs

## Grouped time series

A \alert{\textbf{grouped time series}} is a collection of time series that can be grouped together in a number of non-hierarchical ways.

\vspace*{-0.2cm}\begin{center}
\begin{minipage}{9.2cm}
\begin{block}{}
\begin{tikzpicture}[level distance=1.5cm]
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,outer sep=0pt, fill=red!15]
\tikzstyle{level 1}=[sibling distance=23mm,set style={{every node}+=[fill=blue!15]},level distance=1cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}, level distance=0.9cm]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AX}}
   child {node {AY}}
 }
 child {node {B}
   child {node {BX}}
   child {node {BY}}
 };
\end{tikzpicture}\hspace*{1cm}
\begin{tikzpicture}[level distance=1.5cm]
\tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,outer sep=0pt, fill=red!15]
\tikzstyle{level 1}=[sibling distance=23mm,set style={{every node}+=[fill=blue!15]},level distance=1cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=yellow]}, level distance=0.9cm]
\node{Total}[edge from parent fork down]
 child {node {X}
   child {node {AX}}
   child {node {BX}}
 }
 child {node {Y}
   child {node {AY}}
   child {node {BY}}
 };
\end{tikzpicture}
\end{block}
\end{minipage}
\end{center}

\pause\alert{Examples}\vspace*{-0.4cm}

 * Tourism by state and purpose of travel
 * Retail sales by product groups/sub groups, and by countries/regions

## Hierarchical and grouped time series

\begin{textblock}{8.8}(0.2,1.5)
Almost all collections of time series with linear constraints can be written \rlap{as}
\centerline{\colorbox[RGB]{210,210,210}{$\bY_{t}=\color{blue}\bS\color{red}\bm{b}_{t}$}}
\vspace*{-0.9cm}\begin{itemize}\parskip=0cm\itemsep=0cm
\item $\by_t=$ vector of all series at time $t$
\item $ y_{\text{Total},t}= $ aggregate of all series at time
$t$.
\item $ y_{X,t}= $ value of series $X$ at time $t$.
\item $\color{red}{\bm{b}_t}=$ vector of most disaggregated series at time $t$
\item $\color{blue}{\bS}=$ ``summing matrix'' containing the linear constraints.
\end{itemize}
\end{textblock}

\begin{textblock}{5.7}(11.4,0.1)
\begin{minipage}{4cm}
\begin{block}{}\centering
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
\tikzstyle[level distance=.3cm]
\tikzstyle[sibling distance=12cm]
\tikzstyle{level 1}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=blue!15]}]
\node{Total}[edge from parent fork down]
 child {node {A}
 }
 child {node {B}
 }
 child {node {C}
 };
\end{tikzpicture}
\end{block}
\end{minipage}
\end{textblock}

\begin{textblock}{5.7}(9.4,2.8)\fontsize{14}{15}\sf
\begin{align*}
\bY_{t}&= \begin{pmatrix}
  y_{\text{Total},t}\\
  y_{A,t}\\
  y_{B,t}\\
  y_{C,t}
  \end{pmatrix}  \\
  &= {\color{blue}\underbrace{\begin{pmatrix}
                1 & 1 & 1 \\
                1 & 0 & 0 \\
                0 & 1 & 0\\
                0 & 0 & 1
                \end{pmatrix}}_{\bS}}
     {\color{red}\underbrace{\begin{pmatrix}
       y_{A,t}\\y_{B,t}\\y_{C,t}
       \end{pmatrix}}_{\bm{b}_{t}}}
\end{align*}
\end{textblock}

## Hierarchical time series

\begin{block}{}\hspace*{.6cm}{\centering\small
\begin{tikzpicture}[level distance=1cm]
\tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
\tikzstyle[level distance=.01cm]
\tikzstyle[sibling distance=12cm]
\tikzstyle{level 2}=[sibling distance=10mm,font=\scriptsize,set style={{every node}+=[fill=yellow]}]
\tikzstyle{level 1}=[sibling distance=40mm,font=\footnotesize,set style={{every node}+=[fill=blue!15]}]
\node{Total}[edge from parent fork down]
 child {node {A}
   child {node {AX}}
   child {node {AY}}
   child {node {AZ}}
 }
 child {node {B}
   child {node {BX}}
   child {node {BY}}
   child {node {BZ}}
 }
 child {node {C}
   child {node {CX}}
   child {node {CY}}
   child {node {CZ}}
 };
\end{tikzpicture}}
\end{block}\vspace*{0.1cm}\pause\fontsize{8}{8}\sf

\hbox{$\by_{t}= \begin{pmatrix}
    y_{\text{Total},t}\\
    y_{A,t}\\
    y_{B,t}\\
    y_{C,t}\\
    y_{AX,t}\\
    y_{AY,t}\\
    y_{AZ,t}\\
    y_{BX,t}\\
    y_{BY,t}\\
    y_{BZ,t}\\
    y_{CX,t}\\
    y_{CY,t}\\
    y_{CZ,t}\end{pmatrix}=
    {\color{red}{\begin{pmatrix}
                1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1\\
                1 & 1 & 1 & 0 & 0 & 0 & 0 & 0 & 0\\
                0 & 0 & 0 & 1 & 1 & 1 & 0 & 0 & 0\\
                0 & 0 & 0 & 0 & 0 & 0 & 1 & 1 & 1\\
                1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
                0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
                0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0\\
                0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0\\
                0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0\\
                0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0\\
                0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0\\
                0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0\\
                0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\
             \end{pmatrix}}}{\color{blue}{\begin{pmatrix}
    y_{AX,t}\\
    y_{AY,t}\\
    y_{AZ,t}\\
    y_{BX,t}\\
    y_{BY,t}\\
    y_{BZ,t}\\
    y_{CX,t}\\
    y_{CY,t}\\
    y_{CZ,t}\end{pmatrix}}}$}

\only<3>{\begin{textblock}{3}(10.5,8)\fontsize{14}{15}\sf\colorbox[gray]{.8}{$\by_{t}=\color{red}\bS\color{blue}\bm{b}_{t}$}\end{textblock}}

## Grouped data

\begin{block}{}
\begin{center}\small
\tikzstyle{every node}=[inner sep=2pt]
\begin{tikzpicture}
    \matrix[ampersand replacement=\&,column sep=0.3cm] {
        \node[ellipse,draw,fill=yellow,font=\scriptsize,distance=1cm] {AX};~ \&
        \node[ellipse,draw,fill=yellow,font=\scriptsize] {AY};~ \&
        \node[ellipse,draw,fill=blue!15] {A}; \\[0.3cm]
        \node[ellipse,draw,fill=yellow,font=\scriptsize] {BX};~ \&
        \node[ellipse,draw,fill=yellow,font=\scriptsize] {BY};~ \&
        \node[ellipse,draw,fill=blue!15] {B}; \\[0.3cm]
        \node[ellipse,draw,fill=blue!15] {X};~ \&
        \node[ellipse,draw,fill=blue!15] {Y};~ \&
        \node[ellipse,draw,fill=red!15] {Total}; \\
};
\end{tikzpicture}
\end{center}
\end{block}\pause\fontsize{10}{11}\sf

\hbox{$\by_{t}= \begin{pmatrix}
    y_{\text{Total},t}\\
    y_{A,t}\\
    y_{B,t}\\
    y_{X,t}\\
    y_{Y,t}\\
    y_{AX,t}\\
    y_{AY,t}\\
    y_{BX,t}\\
    y_{BY,t}
    \end{pmatrix}=
    \color{red}\begin{pmatrix}
                1 & 1 & 1 & 1 \\
                1 & 1 & 0 & 0 \\
                0 & 0 & 1 & 1 \\
                1 & 0 & 1 & 0 \\
                0 & 1 & 0 & 1 \\
                1 & 0 & 0 & 0 \\
                0 & 1 & 0 & 0 \\
                0 & 0 & 1 & 0 \\
                0 & 0 & 0 & 1
             \end{pmatrix}
    \color{blue}\begin{pmatrix}
    y_{AX,t}\\
    y_{AY,t}\\
    y_{BX,t}\\
    y_{BY,t}
    \end{pmatrix}$}

\vspace*{-1cm}

\only<3>{\begin{textblock}{3}(10.5,8)\fontsize{14}{15}\sf\colorbox[gray]{.8}{$\by_{t}=\color{red}\bS\color{blue}\bm{b}_{t}$}\end{textblock}}

## The hierarchical forecasting problem

* We want forecasts at all levels of aggregation.
* If we model and forecast each series independently, the forecasts will almost certainly not add up.
* We need to impose constraints on the forecasts to ensure they are "coherent".
* We need to do this in a way that is computationally efficient.


## Constraint notation

\begin{textblock}{5.7}(11.4,0.1)
\begin{minipage}{4cm}
\begin{block}{}\centering
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
\tikzstyle[level distance=.3cm]
\tikzstyle[sibling distance=12cm]
\tikzstyle{level 1}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=blue!15]}]
\node{Total}[edge from parent fork down]
 child {node {A}
 }
 child {node {B}
 }
 child {node {C}
 };
\end{tikzpicture}
\end{block}
\end{minipage}
\end{textblock}

\begin{textblock}{6}(0.5,1.5)
\begin{block}{Aggregation matrix}\vspace*{-0.6cm}
\begin{align*}
\bY_{t} & =\color{blue}\bS\color{red}\bm{b}_{t} \\[0.3cm]
\begin{pmatrix}
   \textcolor{DarkYellow}{y_{\text{Total},t}}\\
   \textcolor{red}{y_{A,t}}\\
   \textcolor{red}{y_{B,t}}\\
   \textcolor{red}{y_{C,t}}
  \end{pmatrix}
  &= {\color{blue}\begin{pmatrix}
                \textcolor{DarkYellow}1 & \textcolor{DarkYellow}1 & \textcolor{DarkYellow}1 \\
                1 & 0 & 0 \\
                0 & 1 & 0\\
                0 & 0 & 1
                \end{pmatrix}}
     {\color{red}\begin{pmatrix}
       y_{A,t}\\y_{B,t}\\y_{C,t}
       \end{pmatrix}} \\[0.2cm]
  \begin{pmatrix}\textcolor{DarkYellow}{\bm{a}_t}\\\textcolor{red}{\bm{b}_t}\end{pmatrix}
   & = \begin{pmatrix}\textcolor{DarkYellow}{\bm{A}}\\\bm{I}_{n_b}\end{pmatrix}\textcolor{red}{\bm{b}_t}
\end{align*}
\end{block}
\end{textblock}

\begin{textblock}{7}(8.1,3.5)
\begin{block}{Constraint matrix}\vspace*{-0.6cm}
\begin{align*}
\bm{C} \bY_t & = \bm{0} \\
\text{where}\qquad
\bm{C} & =  \begin{bmatrix} 1 & -1 & -1 & -1 \end{bmatrix} \\
  & = \begin{bmatrix} \bm{I}_{n_a} & -\textcolor{DarkYellow}{\bm{A}} \end{bmatrix}
\end{align*}
\end{block}
\end{textblock}


## Constraint notation
\vspace*{0.2cm}

\begin{block}{Aggregation matrix $\bm{A}$}
$$\bm{y}_t
    = \begin{bmatrix}\bm{a}_t\\\bm{b}_t\end{bmatrix}
    = \begin{bmatrix}\bm{A}\\\bm{I}_{n_b}\end{bmatrix}\bm{b}_t
    = \bm{S}\bm{b}_t
$$
\end{block}\pause

\begin{block}{Constraint matrix $\bm{C}$}
\centerline{$\bm{C}\bm{y}_t = \bm{0}$}
\end{block}\vspace*{-0.3cm}

* Constraint matrix approach more general & more parsimonious.
* $\bm{C} = [\bm{I}_{n_a} ~~~ {-\bm{A}}]$.
* $\bm{S}$, $\bm{A}$ and $\bm{C}$ may contain any real values (not just 0s and 1s).



## The coherent subspace

\begin{textblock}{9}(.2,1)\fontsize{13}{13}\sf
\begin{block}{Coherent subspace}
$n_b$-dimensional linear subspace $\mathfrak{s}\subset \mathbb{\chi}^n$ for which linear constraints hold for all $\bm{y}\in\mathfrak{s}$.
\end{block}\vspace*{-0.3cm}
\begin{block}{Hierarchical time series}
An $n$-dimensional multivariate time series such that $\bm{y}_t\in\mathfrak{s}\quad\forall t$.
\end{block}\vspace*{-0.3cm}
\begin{block}{Coherent point forecasts}
$\tilde{\bm{y}}_{t+h|t}$ is \emph{coherent} if $\tilde{\bm{y}}_{t+h|t} \in \mathfrak{s}$.
\end{block}\vspace*{-0.2cm}
\end{textblock}
\only<2-3>{\begin{textblock}{7.5}(.2,6.75)\fontsize{13}{13}\sf
\begin{alertblock}{Base forecasts}
Let $\hat{\bm{y}}_{t+h|t}$ be vector of \emph{incoherent} initial $h$-step forecasts.$\phantom{y_{t|h}}$
\end{alertblock}
\end{textblock}}
\only<3>{\begin{textblock}{7.5}(8.3,6.75)\fontsize{13}{13}\sf
\begin{alertblock}{Reconciled forecasts}
Let $\psi$ be a mapping, $\psi:\mathbb{\chi}^n\rightarrow\mathfrak{s}$.  $\tilde{\bm{y}}_{t+h|t}=\psi(\hat{\bm{y}}_{t+h|t})$ ``reconciles'' $\hat{\bm{y}}_{t+h|t}$.
\end{alertblock}
\end{textblock}}

\placefig{9.4}{.0}{width=6.6cm}{3D_hierarchy}
\begin{textblock}{3}(11.4,5.6)\fontsize{13}{13}\sf
\begin{block}{}
\centerline{$y_{Tot} = y_A + y_B$}
\end{block}
\end{textblock}

## Linear projection reconciliation

\only<1>{\placefig{9.5}{1.5}{width=6.2cm}{InsampDir_2_George}}
\only<2>{\placefig{9.5}{1.5}{width=6.2cm}{OrthProj_George}}
\only<3->{\placefig{9.5}{1.5}{width=6.2cm}{ObliqProj_George}}

\begin{textblock}{9}(.5,1.5)
  \begin{itemize}\tightlist
  \item $R$ is the most likely direction of deviations from $\mathfrak{s}$.
  \only<1->{\item Grey: potential base forecasts}
  \only<2->{\item Red: reconciled forecasts}
  \only<2->{\item Orthogonal projections (i.e., OLS) lead to smallest possible adjustments of base forecasts.}
  \only<3->{\item Oblique projections (i.e., MinT) give reconciled forecasts with smallest variance.}
  \end{itemize}
\end{textblock}

\only<2->{
  \begin{textblock}{4.6}(11.2,7.6)\fontsize{13}{14}\sf
  \begin{block}{}
  \only<2>{Orthogonal \rlap{projection}}
  \only<3>{Oblique projection}
  \end{block}
  \end{textblock}
}

## Linear projection reconciliation
\fontsize{14}{16}\sf
\vspace*{0.2cm}\begin{alertblock}{}
\centerline{$\tilde{\bm{y}}_{t+h|t}= \psi(\hat{\bm{y}}_{t+h|t}) = \bm{M}\hat{\bm{y}}_{t+h|t}$}
\end{alertblock}\vspace*{-0.5cm}

* $\bm{M}$ is a projection onto $\mathfrak{s}$ if and only if $\bm{M}\bm{y}=\bm{y}$ for all $\bm{y}\in\mathfrak{s}$.
* Coherent base forecasts are unchanged since $\bm{M}\hat{\bm{y}}=\hat{\bm{y}}$
* If $\hat{\bm{y}}$ is unbiased, then $\tilde{\bm{y}}$ is also unbiased since
$$
  \E(\tilde{\bm{y}}_{t+h|t}) = \E(\bm{M}\hat{\bm{y}}_{t+h|t}) = \bm{M} \E(\hat{\bm{y}}_{t+h|t}) = \E(\hat{\bm{y}}_{t+h|t}),
$$
and unbiased estimates must lie on $\mathfrak{s}$.
* If $\bm{S}$ forms a basis set for $\mathfrak{s}$, then projections are of the form $\bm{M} = \bS(\bS'\bm{\Psi}\bS)^{-1}\bS'\bm{\Psi}$ where $\bm{\Psi}$ is a positive definite matrix.
* The projection is orthogonal if and only if $\bm{M}'=\bm{M}$. Equivalently, $\bm{\Psi}=\bm{I}$.

\vspace*{10cm}

## Linear projection reconciliation

\vspace*{0.2cm}\begin{alertblock}{}
\centerline{$\tilde{\bm{y}}_{t+h|t}= \psi(\hat{\bm{y}}_{t+h|t}) = \bm{M}\hat{\bm{y}}_{t+h|t}$}\vspace{0.1cm}

\centerline{where\quad $\bm{M} = \bS(\bS'\bm{\Psi}\bS)^{-1}\bS'\bm{\Psi} = \bm{I} - \bm{\Psi}\bm{C}'(\bm{C}\bm{\Psi}\bm{C}')^{-1}\bm{C}
$}
\end{alertblock}\vspace*{.01cm}
\begin{block}{}\vspace*{-0.6cm}
\begin{align*}
&\text{OLS:}  && \bm{\Psi}=\bm{I} \\
&\text{WLS:}  && \bm{\Psi}=\text{diagonal} \\
&\text{MinT:} && \bm{\Psi}=\bm{W}_h
\end{align*}
\end{block}\vspace*{-0.3cm}

* $\bm{W}_h = \var[\by_{T+h} - \hat{\by}_{T+h|T} \mid \by_1,\dots,\by_T]$ is the covariance matrix of the base forecast errors.
* $\bm{V}_h = \var[\by_{T+h} - \tilde{\by}_{T+h|T}  \mid \by_1,\dots,\by_T])  = \bm{M}\bm{W}_h\bm{M}'$ is the covariance matrix of the reconciled forecast errors.

\vspace*{10cm}

## Minimum trace reconciliation

\vspace*{0.2cm}\begin{alertblock}{Minimum trace (MinT) reconciliation}
If $\bm{M}$ is a projection, then trace of $\bm{V}_h$ is minimized when $\bm{\Psi} = \bm{W}_h$, so that
\centerline{$\bm{M} = \bS(\bS'\bm{W}_h^{-1}\bS)^{-1}\bS'\bm{W}_h^{-1} =\bm{I} - \bm{W}_h\bm{C}'(\bm{C}\bm{W}_h\bm{C}')^{-1}\bm{C}$}
\end{alertblock}
\begin{block}{}
\centerline{$\displaystyle\textcolor{red}{\tilde{\by}_{T+h|T}}
= \bm{M} ~ \textcolor{blue}{\hat{\by}_{T+h|T}}$}
\end{block}\vspace*{-0.2cm}
\centerline{\hspace*{1.4cm}\textcolor{red}{Reconciled forecasts}\hfill\textcolor{blue}{Base forecasts}\hspace*{2.9cm}}\vspace*{-0.2cm}

* Trace of $\bm{V}_h$ is sum of forecast variances.
* MinT is $L_2$ optimal amongst linear unbiased forecasts.
* How to estimate $\bm{W}_h = \var[\by_{T+h} - \hat{\by}_{T+h|T} \mid \by_1,\dots,\by_T]$?

## Linear projections

\begin{textblock}{5}(7,-0.2)
\begin{block}{}
\centerline{$\tilde{\by}_{T+h|T}=\bS\bG\hat{\by}_{T+h|T}$}
\end{block}
\end{textblock}

\begin{textblock}{9.4}(.5,1.2)
\begin{alertblock}{Reconciliation method \hspace*{0.5cm} $\bG$}
\begin{tabular}{ll}
  OLS             & $(\bS'\bS)^{-1}\bS'$ \\
  WLS(var)        & $(\bS'\bm{\Lambda}_v\bS)^{-1}\bS'\bm{\Lambda}_v$ \\
  WLS(struct)     & $(\bS'\bm{\Lambda}_s\bS)^{-1}\bS'\bm{\Lambda}_s$ \\
  MinT(sample)    & $(\bS'\hat{\bm{W}}_{\text{sam}}^{-1}\bS)^{-1}\bS' \hat{\bm{W}}_{\text{sam}}^{-1}$  \\
  MinT(shrink)\hspace*{2cm}    & $(\bS'\hat{\bm{W}}_{\text{shr}}^{-1}\bS)^{-1}\bS' \hat{\bm{W}}_{\text{shr}}^{-1}$  \\
\end{tabular}
\end{alertblock}
\end{textblock}
\begin{textblock}{15}(.2,5.7)\fontsize{13}{15}\sf
\begin{itemize}\parskip=0cm
\item $\bm{\Lambda}_v = \text{diag}(\bm{W}_1)^{-1}$
\item $\hat{\bm{W}}_{\text{sam}}$ is sample estimate of the residual covariance matrix
\item $\hat{\bm{W}}_{\text{shr}}$ is shrinkage estimator $\tau \text{diag}(\hat{\bm{W}}_{\text{sam}})+(1-\tau)\hat{\bm{W}}_{\text{sam}}$\\ where $\tau$ selected optimally.
\item Still need a good estimate of $\bm{W}_h$ for forecast variance.
\end{itemize}
\end{textblock}
\begin{textblock}{14}(4.5,5.7)\fontsize{13}{15}\sf
\begin{itemize}\tightlist
\item $\bm{\Lambda}_s = \text{diag}(\bS\bm{1})^{-1}$
\end{itemize}
\end{textblock}
\begin{textblock}{5}(10.3,3.35)
\begin{block}{}
These approximate MinT by assuming $\bm{W}_h = k_h \bm{W}_1$.
\end{block}
\end{textblock}

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r tourismdata00, echo=TRUE}
tourism
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r tourismagg, echo=TRUE}
tourism_agg <- tourism |>
  aggregate_key(state / zone / region, visitors = sum(visitors))
```

```{r tourismagg2, echo=FALSE}
tourism_agg
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r tourismmodels, echo=TRUE}
fit <- tourism_agg |>
  filter(year(month) <= 2015) |>
  model(ets = ETS(visitors))
```

```{r tourismmodels1, echo=FALSE}
fit
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism, echo=TRUE}
fc <- fit |>
  reconcile(
    ols = min_trace(ets, method = "ols"),
    wlsv = min_trace(ets, method = "wls_var"),
    wlss = min_trace(ets, method = "wls_struct"),
    # mint_c = min_trace(ets, method="mint_cov"),
    mint_s = min_trace(ets, method = "mint_shrink"),
  ) |>
  forecast(h = "2 years")
```

```{r fctourism1, echo=FALSE}
fc
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism2, dependson='fctourism', echo=TRUE, fig.width=12, fig.height=4.5}
fc |>
  filter(is_aggregated(state)) |>
  autoplot(filter(tourism_agg, year(month) > 2012), level = NULL)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism3, dependson='fctourism', echo=TRUE, fig.width=12, fig.height=4.5}
fc |>
  filter(state == "NSW" & is_aggregated(zone)) |>
  autoplot(filter(tourism_agg, year(month) > 2012), level = NULL)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism4, dependson='fctourism', echo=TRUE, fig.width=12, fig.height=4.5}
fc |>
  filter(region == "Melbourne") |>
  autoplot(filter(tourism_agg, year(month) > 2012), level = NULL)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism5, dependson='fctourism', echo=TRUE, fig.width=12, fig.height=4.5}
fc |>
  filter(region == "Snowy Mountains") |>
  autoplot(filter(tourism_agg, year(month) > 2012), level = NULL)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fctourism6, dependson='fctourism', echo=TRUE, fig.width=12, fig.height=4.5}
fc |>
  filter(region == "Barossa") |>
  autoplot(filter(tourism_agg, year(month) > 2012), level = NULL)
```

## Performance evaluation

\vspace*{0.2cm}\begin{block}{}
\centerline{$\text{MASE} = \text{mean}(|q_{j}|)$}
$$
  q_{j} = \frac{e_{j}\phantom{^2}}
 {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T |y_{t}-y_{t-m}|}
$$
\end{block}

* $y_t=$ observation for period $t$
* $e_{j}=$ forecast error for forecast horizon $j$
* $T=$ size of training set
* $m = 12$

## Performance evaluation

\vspace*{0.2cm}\begin{block}{}
\centerline{$\text{RMSSE} = \sqrt{\text{mean}(q_{j}^2)}$}
$$
  q^2_{j} = \frac{ e^2_{j}}
 {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T (y_{t}-y_{t-m})^2}
$$
\end{block}

* $y_t=$ observation for period $t$
* $e_{j}=$ forecast error for forecast horizon $j$
* $T=$ size of training set
* $m = 12$

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fcaccuracy, dependson='fctourismcomb', echo=TRUE}
fc |>
  accuracy(tourism_agg, measures = list(mase = MASE, rmsse = RMSSE))
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```{r fcaccuracy2, dependson='fctourismcomb', echo=TRUE}
fc |>
  accuracy(tourism_agg, measures = list(mase = MASE, rmsse = RMSSE)) |>
  group_by(.model) |>
  summarise(mase = mean(mase), rmsse = sqrt(mean(rmsse^2))) |>
  arrange(rmsse)
```

\only<2>{\begin{textblock}{7}(7,5)
\begin{block}{}\fontsize{13}{14}\sf
\begin{itemize}\tightlist
\item Overall, every reconciliation method is better than the base ETS forecasts.
\end{itemize}
\end{block}
\end{textblock}}

## Example: Australian tourism
\fontsize{8}{7}\sf

```{r fcaccuracy3, dependson='fctourismcomb', echo=FALSE}
fc |>
  accuracy(tourism_agg,
    measures = list(mase = MASE, rmsse = RMSSE)
  ) |>
  mutate(
    level = case_when(
      is_aggregated(state) ~ "National",
      is_aggregated(zone) ~ "State",
      is_aggregated(region) ~ "Zone",
      TRUE ~ "Region"
    ),
    level = factor(level, levels = c("National", "State", "Zone", "Region"))
  ) |>
  group_by(.model, level) |>
  summarise(mase = mean(mase), rmsse = sqrt(mean(rmsse^2))) |>
  arrange(level, rmsse)
```

\only<2>{\begin{textblock}{7}(7,5)
\begin{block}{}\fontsize{13}{14}\sf
\begin{itemize}\tightlist
\item OLS is best for all levels except national.
\item Improvements due to reconciliation are greater at lower levels.
\end{itemize}
\end{block}
\end{textblock}}

## Mean square error bounds

\begin{textblock}{6.4}(9,-0.1)\begin{block}{}
\citet{htsgeometry}
\end{block}\end{textblock}

\vspace*{0.2cm}\begin{alertblock}{Distance reducing property}
Let $\|\bm{u}\|_{\bm{\Psi}} = \bm{u}'\bm{\Psi}\bm{u}$. Then
  \centerline{$\|\bm{y}_{t+h}-\tilde{\bm{y}}_{t+h|t}\|_{\bm{\Psi}}\le\|\bm{y}_{t+h}-\hat{\bm{y}}_{t+h|t}\|_{\bm{\Psi}}$}
\end{alertblock}

 * $\bm{\Psi}$-projection is guaranteed to improve forecast accuracy over base forecasts *using this distance measure*.
 * Distance reduction holds for any realisation and any forecast.
 * OLS reconciliation minimizes Euclidean distance.

## Mean square error bounds

\begin{textblock}{6.4}(9,-0.1)\begin{block}{}
\citet{wickramasuriya2021properties}
\end{block}\end{textblock}

\begin{block}{}\vspace*{-0.6cm}
\begin{align*}
\|\bm{y}_{t+h} - \tilde{\bm{y}}_{t+h}\|_2^2
 &= \|\bm{M}(\bm{y}_{t+h} - \hat{\bm{y}}_{t+h})\|_2^2 \\
 &\le \|\bm{M}\|_2^2 \|\bm{y}_{t+h} - \hat{\bm{y}}_{t+h}\|_2^2 \\
 & = \sigma_{\text{max}}^2\|\bm{y}_{t+h} - \hat{\bm{y}}_{t+h}\|_2^2
\end{align*}
\end{block}

 * $\sigma_{\text{max}}$ is the largest eigenvalue of $\bm{M}$
 * $\sigma_{\text{max}}\ge1$ as $\bm{M}$ is a projection matrix.
 * Every projection reconciliation is better than base forecasts using Euclidean distance.

## Mean square error bounds

\begin{textblock}{6.4}(9,-0.1)\begin{block}{}
\citet{wickramasuriya2021properties}
\end{block}\end{textblock}
\vspace*{0.2cm}\begin{block}{}\vspace*{-0.6cm}
\begin{align*}
    & \text{tr}\Big(\E[\bm{y}_{t+h} - \tilde{\bm{y}}^{\text{MinT}}_{t+h|t}]'[\bm{y}_{t+h} - \tilde{\bm{y}}^{\text{MinT}}_{t+h|t}]\Big) \\
\le~ & \text{tr}\Big(\E[\bm{y}_{t+h} - \tilde{\bm{y}}^{\text{OLS}}_{t+h|t}]'[\bm{y}_{t+h} - \tilde{\bm{y}}^{\text{OLS}}_{t+h|t}]\Big) \\
\le~ & \text{tr}\Big(\E[\bm{y}_{t+h} - \hat{\bm{y}}_{t+h|t}]'[\bm{y}_{t+h} - \hat{\bm{y}}_{t+h|t}]\Big)
\end{align*}
\end{block}

Using sums of variances:

* MinT reconciliation is better than OLS reconciliation
* OLS reconciliation is better than base forecasts

## Non-negative forecasts
\fontsize{14}{15}\sf
\vspace*{0.1cm}\begin{alertblock}{}
\centerline{$\min_{\bm{G}_h}\text{tr}\Big(\E[\bm{y}_{t+h} - \bm{S}\bm{G}_h\hat{\bm{y}}_{t+h|t}]'[\bm{y}_{t+h} - \bm{S}\bm{G}_h\hat{\bm{y}}_{t+h|t}]\Big)$}
\centerline{such that $\bm{b}_{t+h|t} = \bm{G}_h\hat{\bm{y}}_{t+h|t} \ge 0$}
\end{alertblock}\pause\vspace*{-0.2cm}

\begin{block}{Solve via quadratic programming:}
$$
  \text{min}_{\bm{b}} \big[\bm{b}'\bS'\bm{W}_h^{-1}\bS\bm{b} - 2 \bm{b}'\bS'\bm{W}_h^{-1}\hat{\bm{y}}_{T+h|T}\big] \quad \text{s.t.~~} \bm{b} \ge 0
$$
\rightline{\citep{nonnegmint}}
\end{block}\pause\vspace*{-0.1cm}

### Set-negative-to-zero heuristic solution
  * Negative reconciled forecasts at bottom level set to zero
  * Remaining forecasts computed via aggregation

\rightline{\citep{di2023spatio}}


## Reconciliation and regularization

@Mishchenko2019:\newline Optimize all forecasts with an incoherence penalty
$$
\text{min}_{\hat{\bm{y}}_{t}} \sum_{t=1}^T \|\bm{y}_t - \hat{\bm{y}}_t\|_2 + \lambda \sum_{t=1}^T \|\hat{\bm{y}}_t - \bm{S}_t \hat{\bm{b}}_t\|_2
$$
\pause


@Shiratori2020:\newline Optimize bottom level forecasts with an incoherence penalty
$$
\text{min}_{\hat{\bm{b}}_{t}} \sum_{t=1}^T \|\bm{b}_t - \hat{\bm{b}}_t\|_2 + \sum_{t=1}^T  \|\bm{\Lambda}(\bm{a}_t - \bm{A}_t \hat{\bm{b}}_t)\|_2
$$


\nocite{hierarchical,fasthts, mint, hfreview, htsgeometry, nonnegmint, Di_FonGir2022b, lhf}


```{r}
#| cache: false
library(knitr)
library(kableExtra)
```


# Improving univariate forecasts

## Temporal reconciliation: quarterly data

\begin{textblock}{7}(1,1.9)
\begin{tikzpicture}
  \tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15,font=\small]
  \tikzstyle[level distance=.1cm]
  \tikzstyle[sibling distance=3cm]
  \tikzstyle{level 1}=[sibling distance=18mm,set style={{every node}+=[fill=yellow]}]
  \node{Annual}[edge from parent fork down]
  child {node {Q$_1$}}
  child {node {Q$_2$}}
  child {node {Q$_3$}}
  child {node {Q$_4$}};
\end{tikzpicture}
\end{textblock}

\only<3>{\begin{textblock}{8.5}(8.3,1.1)\fontsize{14}{17}\sf
$$\bm{y}_\tau =
  \begin{bmatrix}
    x_{\tau}^{[4]}  \\[0.2cm]
    x_{\tau,1}^{[1]} \\[0.2cm]
    x_{\tau,2}^{[1]} \\[0.2cm]
    x_{\tau,3}^{[1]} \\[0.2cm]
    x_{\tau,4}^{[1]}
  \end{bmatrix}
  \qquad
  \bm{S} = \begin{bmatrix}
    1 & 1 & 1 & 1 \\[0.2cm]
    1 & 0 & 0 & 0 \\[0.2cm]
    0 & 1 & 0 & 0 \\[0.2cm]
    0 & 0 & 1 & 0 \\[0.2cm]
    0 & 0 & 0 & 1
  \end{bmatrix}
$$
\end{textblock}}

\only<2->{\begin{textblock}{8}(.3,5.7)
  \begin{alertblock}{}
    \begin{itemize}
      \item[\color{white}\ding{229}] Forecast series at each available frequency.
      \item[\color{white}\ding{229}] Optimally combine forecasts within the same year.
    \end{itemize}
  \end{alertblock}
\end{textblock}}

\only<3->{\begin{textblock}{5}(10.3,7.9)
\fontsize{11}{12}\sf
$\tau=$ index of largest temporal aggregation level.
\end{textblock}}


## Temporal reconciliation: quarterly data

\begin{textblock}{7}(0.1,1.9)
\begin{tikzpicture}
  \tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15,font=\small]
  \tikzstyle[level distance=.1cm]
  \tikzstyle[sibling distance=3cm]
  \tikzstyle{level 1}=[sibling distance=42mm,set style={{every node}+=[fill=blue!15]}]
  \tikzstyle{level 2}=[sibling distance=18mm,set style={{every node}+=[fill=yellow]}]
  \node{Annual}[edge from parent fork down]
  child {node {Semi-Annual$_1$}
      child {node {Q$_1$}}
      child {node {Q$_2$}}
    }
  child {node {Semi-Annual$_2$}
      child {node {Q$_3$}}
      child {node {Q$_4$}}
    };
\end{tikzpicture}
\end{textblock}

\begin{textblock}{8.5}(8.3,1.1)\fontsize{14}{17}\sf
$$\bm{y}_\tau =
  \begin{bmatrix}
    x_{\tau}^{[4]}  \\[0.2cm]
    x_{\tau,1}^{[2]} \\[0.2cm]
    x_{\tau,2}^{[2]} \\[0.2cm]
    x_{\tau,1}^{[1]} \\[0.2cm]
    x_{\tau,2}^{[1]} \\[0.2cm]
    x_{\tau,3}^{[1]} \\[0.2cm]
    x_{\tau,4}^{[1]}
  \end{bmatrix}
  \qquad
  \bm{S} = \begin{bmatrix}
    1 & 1 & 1 & 1 \\[0.2cm]
    1 & 1 & 0 & 0 \\[0.2cm]
    0 & 0 & 1 & 1 \\[0.2cm]
    1 & 0 & 0 & 0 \\[0.2cm]
    0 & 1 & 0 & 0 \\[0.2cm]
    0 & 0 & 1 & 0 \\[0.2cm]
    0 & 0 & 0 & 1
  \end{bmatrix}
$$
\end{textblock}

\begin{textblock}{8}(.3,5.7)
  \begin{alertblock}{}
    \begin{itemize}
      \item[\color{white}\ding{229}] Forecast series at each available frequency.
      \item[\color{white}\ding{229}] Optimally combine forecasts within the same year.
    \end{itemize}
  \end{alertblock}
\end{textblock}

\begin{textblock}{5}(10.3,7.9)
\fontsize{11}{12}\sf
$\tau=$ index of largest temporal aggregation level.
\end{textblock}

## Temporal reconciliation: monthly data

\only<1>{\begin{tikzpicture}
  \tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15,font=\small]
  \tikzstyle[level distance=.1cm]
  \tikzstyle[sibling distance=7cm]
  \tikzstyle{level 1}=[sibling distance=72mm,set style={{every node}+=[fill=blue!15]}]
  \tikzstyle{level 2}=[sibling distance=36mm,set style={{every node}+=[fill=yellow]}]
  \tikzstyle{level 3}=[sibling distance=12mm,font=\scriptsize,set style={{every node}+=[fill=green]}]
  \node{Annual}[edge from parent fork down]
  child {node {Semi-Annual$_1$}
      child {node {Q$_1$}
          child {node {\scriptsize M$_1$}}
          child {node {\scriptsize M$_2$}}
          child {node {\scriptsize M$_3$}}
        }
      child {node {Q$_2$}
          child {node {\scriptsize M$_4$}}
          child {node {\scriptsize M$_5$}}
          child {node {\scriptsize M$_6$}}
        }
    }
  child {node {Semi-Annual$_2$}
      child {node {Q$_3$}
          child {node {\scriptsize M$_7$}}
          child {node {\scriptsize M$_8$}}
          child {node {\scriptsize M$_9$}}
        }
      child {node {Q$_4$}
          child {node {\scriptsize M$_{10}$}}
          child {node {\scriptsize M$_{11}$}}
          child {node {\scriptsize M$_{12}$}}
        }
    };
\end{tikzpicture}}
\only<2->{\begin{tikzpicture}
  \tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15,font=\small]
  \tikzstyle[level distance=.1cm]
  \tikzstyle[sibling distance=7cm]
  \tikzstyle{level 1}=[sibling distance=48mm,set style={{every node}+=[fill=blue!15]}]
  \tikzstyle{level 2}=[sibling distance=24mm,set style={{every node}+=[fill=yellow]}]
  \tikzstyle{level 3}=[sibling distance=12mm,set style={{every node}+=[fill=green]}]
  \node{Annual}[edge from parent fork down]
  child {node {FourM$_1$}
      child {node {BiM$_1$}
          child {node {\scriptsize M$_1$}}
          child {node {\scriptsize M$_2$}}
        }
      child {node {BiM$_2$}
          child {node {\scriptsize M$_3$}}
          child {node {\scriptsize M$_4$}}
        }
    }
  child {node {FourM$_2$}
      child {node {BiM$_3$}
          child {node {\scriptsize M$_5$}}
          child {node {\scriptsize M$_6$}}
        }
      child {node {BiM$_4$}
          child {node {\scriptsize M$_7$}}
          child {node {\scriptsize M$_8$}}
        }
    }
  child {node {FourM$_3$}
      child {node {BiM$_5$}
          child {node {\scriptsize M$_9$}}
          child {node {\scriptsize M$_{10}$}}
        }
      child {node {BiM$_6$}
          child {node {\scriptsize M$_{11}$}}
          child {node {\scriptsize M$_{12}$}}
        }
    };
\end{tikzpicture}}\pause

\begin{textblock}{14}(1,6.7)
  \begin{alertblock}{}
    \begin{itemize}
      \item[\color{white}\ding{229}] Forecast series at each available frequency.
      \item[\color{white}\ding{229}] Optimally combine forecasts within the same year.
    \end{itemize}
  \end{alertblock}
\end{textblock}

## Temporal reconciliation: monthly data
\fontsize{11}{11}\sf

$$
  \bm{y}_\tau=\begin{bmatrix}
    x_{\tau}^{[12]}     \\[0.2cm]
    \bm{x}_{\tau}^{[6]} \\[0.2cm]
    \bm{x}_{\tau}^{[4]} \\[0.2cm]
    \bm{x}_{\tau}^{[3]} \\[0.2cm]
    \bm{x}_\tau^{[2]}   \\[0.2cm]
    \bm{x}_\tau^{[1]}
  \end{bmatrix}
  \qquad
  \bm{S} = \begin{bmatrix}
    1                & 1 & 1 & 1 & 1~~~1~~~1~~~1 & 1 & 1 & 1 & 1 \\
    1                & 1 & 1 & 1 & 1~~~1~~~0~~~0 & 0 & 0 & 0 & 0 \\
    0                & 0 & 0 & 0 & 0~~~0~~~1~~~1 & 1 & 1 & 1 & 1 \\
    1                & 1 & 1 & 1 & 0~~~0~~~0~~~0 & 0 & 0 & 0 & 0 \\
    0                & 0 & 0 & 0 & 1~~~1~~~1~~~1 & 0 & 0 & 0 & 0 \\
    0                & 0 & 0 & 0 & 0~~~0~~~0~~~0 & 1 & 1 & 1 & 1 \\
    1                & 1 & 1 & 0 & 0~~~0~~~0~~~0 & 0 & 0 & 0 & 0 \\
                     &   &   &   & \vdots        &   &   &   &   \\
    0                & 0 & 0 & 0 & 0~~~0~~~0~~~0 & 0 & 1 & 1 & 1 \\
    1                & 1 & 0 & 0 & 0~~~0~~~0~~~0 & 0 & 0 & 0 & 0 \\
                     &   &   &   & \vdots        &   &   &   &   \\
    0                & 0 & 0 & 0 & 0~~~0~~~0~~~0 & 0 & 0 & 1 & 1 \\[0.2cm]
    \phantom{\vdots} &   &   &   & \bm{I}_{12}   &   &   &   &
  \end{bmatrix}
$$

## Temporal reconciliation
\fontsize{14}{15}\sf

For a time series  $y_1,\dots,y_T$, observed at frequency $m$:
\begin{alertblock}{}\vspace*{-0.1cm}
$$
  x_j^{[k]} = \sum_{t = (j-1)k+1}^{jk} y_t\qquad \text{for $j = 1,\dots,\lfloor T/k\rfloor$}
$$
\end{alertblock}
* $k \in \mathcal{K} = \{k_1,\dots,k_p\}$ denote the $p$ factors of $m$ in ascending order, where $k_1=1$ and $k_p=m$
* $x_j^{[1]} = y_t$
* A single unique hierarchy is only possible when there are no coprime pairs in $\mathcal{K}$.
* $M_k=m/k$ is seasonal period of aggregated series.

## Temporal reconciliation
\fontsize{14}{15}\sf\vspace*{-0.5cm}
$$\bm{x}_\tau = \bm{S} \bm{x}_\tau^{[1]}, \qquad \bm{S} = \begin{bmatrix}\bm{A}\\\bm{I}\end{bmatrix}$$
where
$$
\bm{x}_\tau = \begin{bmatrix*}[l]
    {x_\tau^{[k_p]}}\\
    {\bm{x}_\tau^{[k_{p-1}]}}\\
    \quad\vdots\\
    {\bm{x}_\tau^{[k_1]}}\\
  \end{bmatrix*}\qquad
  \bm{x}_\tau^{[k]} = \begin{bmatrix*}[l]
    x_{M_k(\tau-1)+1}^{[k]}\\
    x_{M_k(\tau-1)+2}^{[k]}\\
    \quad\vdots\\
    x_{M_k\tau}^{[k]}
  \end{bmatrix*}\qquad
\bm{A} =
  \begin{bmatrix}
    \bm{1}'_m                                    \\
    \bm{I}_{m/k_{p-1}} \otimes \bm{1}'_{k_{p-1}} \\
    \vdots                                       \\
    \bm{I}_{m/k_{2}} \otimes \bm{1}'_{k_{2}}     \\
  \end{bmatrix}
$$
$\tau$ is time index for most aggregated series,

$k\in \mathcal{K} = \{k_1,\dots,k_p\}$,\quad $k_1=1$,\quad $k_p=m$,\quad $\tau=1,\dots,T/m$.


## Cross-temporal forecast reconciliation

\alert{Structural matrix approach}

* $\bm{S}_{cs}=$ structural cross-sectional matrix
* $\bm{S}_{te}=$ structural temporal matrix
* $\bm{S}_{ct} = \bm{S}_{cs} \otimes \bm{S}_{te}$

\vspace*{-0.6cm}

$$\bm{x}_\tau = \bm{S}_{ct} \bm{b}_\tau,\qquad\text{where}\qquad
\bm{b}_\tau = \begin{bmatrix}
    {\bm{x}_{1,\tau}^{[1]}} \\
    \vdots                  \\
    {\bm{x}_{n,\tau}^{[1]}}
  \end{bmatrix}.
$$

## Cross-temporal forecast reconciliation

\full{Smatrix}

\begin{textblock}{5}(0.5,1.5)
\begin{block}{}
\centerline{$\bm{S}_{cs} \otimes \bm{S}_{te} = \bm{S}_{ct}$}
\end{block}
\end{textblock}

## Cross-temporal forecast reconciliation

\alert{Constraint matrix approach}

* $\bm{C}_{cs}=$ cross-sectional constraint matrix
* $\bm{C}_{te}=$ temporal constraint matrix

\vspace*{-0.6cm}

$$\bm{C}_{ct}\bm{x}_\tau = \bm{0}\qquad\text{where}\qquad
\bm{C}_{ct} = \begin{bmatrix}
    (\bm{0}_{(n_a m\times nk^\ast)} ~~ \bm{I}_m \otimes \bm{C}_{cs})\bm{P}' \\
    \bm{I}_n \otimes \bm{C}_{te}
  \end{bmatrix}
$$

* $k^\ast = \displaystyle\sum_{k \in \mathcal{K}\setminus\{k_1\}} \frac{m}{k}$
* $\bm{P}=$ the permutation matrix such that $\bm{P} \text{vec}(\bm{Y}_{\tau}) = \text{vec}(\bm{Y}_{\tau}')$.


##  Cross-temporal probabilistic forecast reconciliation

\alert{Nonparametric bootstrap}\fontsize{14}{16}\sf

* Simulate future sample paths from all models using bootstrapped residuals, then reconcile them to obtain coherent sample paths.
* Need to generate samples that preserve cross-temporal relationships.
* Draw residual samples of all series at same time from most temporally aggregated level.
* Residuals for other levels obtained using the corresponding time indices.


##  Cross-temporal probabilistic forecast reconciliation
\vspace*{-0.2cm}

```{r}
#| label: bootres
#| output: asis
boot_plot <- function(cols = c("green","blue","red","black")) {
  cat("\\hspace*{-0.8cm}\\begin{tikzpicture}[scale = 0.9, every node/.style={scale=0.9}]\n")
  cat("\\matrix (e1) [matrix of nodes,ampersand replacement=\\&,row sep=0cm,column sep=0cm, nodes= {rectangle, fill=white, inner sep = 1pt, font = {\\fontsize{7}{6}\\selectfont}, minimum width=1em, minimum height=1em,anchor=center}, label={[xshift = 0.5em]above:{\\footnotesize$\\widehat{\\textbf{E}}^{[1]}$}}]\n")
  cat("{\n")
  cat("$T$ \\& ")
  for(j in cols) {
    for(i in seq(4)) {
      cat(paste0("\\pgfuseimage{n",j, "} \\& "))
    }
  }
  cat("\\\\\n")
  cat("$X$ \\& ")
  for(j in cols) {
    for(i in seq(4)) {
      cat(paste0("\\pgfuseimage{n",j, "} \\& "))
    }
  }
  cat("\\\\\n")
  cat("$Y$ \\& ")
  for(j in cols) {
    for(i in seq(4)){
      cat(paste0("\\pgfuseimage{n",j, "} \\& "))
    }
  }
  cat("\\\\\n")
  for(i in seq(16)) {
    cat(paste("\\& ",i))
  }
  cat("\\\\\n")
  cat("};\n")
  cat("\\draw[decorate,thick, decoration={brace, mirror, amplitude=5pt,raise=-1pt}] (e1-4-2.south west) -- (e1-4-17.south east) node[midway, font = {\\fontsize{7}{6}\\selectfont}, yshift = -1em]{Quarterly, $t = 1,\\dots,16$};\n")
  cat("\\matrix (ek) [above= 10mm of e1.north east, anchor=south east, matrix of nodes,ampersand replacement=\\&,row sep=0cm,column sep=0cm, nodes= {rectangle, fill=white, inner sep = 1pt, font = {\\fontsize{7}{6}\\selectfont}, minimum width=1em, minimum height=1em,anchor=center}, label={[xshift = 0.5em]above:{\\footnotesize$\\widehat{\\textbf{E}}^{[2]}$}}]\n")
  cat("{\n")
  cat("$T$ \\& ")
  for(j in cols) {
    for(i in seq(2)){
     cat(paste0("\\pgfuseimage{n",j, "} \\& "))
    }
  }
  cat("\\\\\n")
  cat("$X$ \\& ")
  for(j in cols) {
    for(i in seq(2)){
     cat(paste0("\\pgfuseimage{n",j, "} \\& "))
    }
  }
  cat("\\\\\n")
  cat("$Y$ \\& ")
  for(j in cols) {
    for(i in seq(2)){
     cat(paste0("\\pgfuseimage{n",j, "} \\& "))
    }
  }
  cat("\\\\\n")
  for(i in seq(8)) {
    cat(paste("\\& ",i))
  }
  cat("\\\\\n")
  cat("};\n")
  cat("\\draw[decorate,thick, decoration={brace, mirror, amplitude=5pt,raise=-1pt}] (ek-4-2.south west) -- (ek-4-9.south east) node[midway, font = {\\fontsize{7}{6}\\selectfont}, yshift = -1em]{Semi-annual, $j = 1,\\dots,8$};\n")
  cat("\\matrix (em) [above= 10mm of e1.north west, anchor=south west, matrix of nodes,ampersand replacement=\\&,row sep=0cm,column sep=0cm, nodes= {rectangle, fill=white, inner sep = 1pt, font = {\\fontsize{7}{6}\\selectfont}, minimum width=1em, minimum height=1em,anchor=center}, label={[xshift = 0.5em]above:{\\footnotesize$\\widehat{\\textbf{E}}^{[4]}$}}]\n")
  cat("{\n")
  cat("$T$ \\& ")
  for(j in cols) {
    cat(paste0("\\pgfuseimage{n",j, "}"))
    if(j != cols[4]) cat("\\& ")
  }
  cat("\\\\\n")
  cat("$X$ \\& ")
  for(j in cols) {
    cat(paste0("\\pgfuseimage{n",j, "}"))
    if(j != cols[4]) cat("\\& ")
  }
  cat("\\\\\n")
  cat("$Y$ \\& ")
  for(j in cols) {
    cat(paste0("\\pgfuseimage{n",j, "}"))
    if(j != cols[4]) cat("\\& ")
  }
  cat("\\\\\n")
  for(i in seq(4)) {
    cat(paste("\\& ",i))
  }
  cat("\\\\\n")
  cat("};\n")
  cat("\\draw[decorate,thick, decoration={brace, mirror, amplitude=5pt,raise=-1pt}] (em-4-2.south west) -- (em-4-5.south east) node[midway, font = {\\fontsize{7}{6}\\selectfont}, yshift = -1em]{Annual, $\\tau = 1,\\dots,4$};\n")
  cat("\\end{tikzpicture}\n\n")
}
boot_plot()
```

\begin{textblock}{1.9}(12.9,5.3)
\begin{block}{}\centering
\textcolor[HTML]{7fb97f}{Year 1}\\
\textcolor[HTML]{7fbcff}{Year 2}\\
\textcolor[HTML]{e87f7f}{Year 3}\\
\textcolor[HTML]{7f7f7f}{Year 4}
\end{block}
\end{textblock}

\begin{textblock}{1.9}(12.9,1.6)
\begin{minipage}{1.9cm}
  \begin{block}{}\centering
    \begin{tikzpicture}
      \tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
      \tikzstyle[level distance=.3cm]
      \tikzstyle[sibling distance=12cm]
      \tikzstyle{level 1}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=blue!15]}]
      \node{Total}[edge from parent fork down]
      child {node {X}
        }
      child {node {Y}
        };
    \end{tikzpicture}
  \end{block}
\end{minipage}
\end{textblock}

\begin{textblock}{3}(3.5,1.2)\alert{Data}\end{textblock}


##  Cross-temporal probabilistic forecast reconciliation
\vspace*{-0.2cm}

```{r}
#| label: bootres1
#| output: asis
#| dependson: bootres
boot_plot(c("red","black","green","blue"))
```

\begin{textblock}{1.9}(12.9,5.3)
\begin{block}{}\centering
\textcolor[HTML]{7fb97f}{Year 1}\\
\textcolor[HTML]{7fbcff}{Year 2}\\
\textcolor[HTML]{e87f7f}{Year 3}\\
\textcolor[HTML]{7f7f7f}{Year 4}
\end{block}
\end{textblock}

\begin{textblock}{1.9}(12.9,1.6)
\begin{minipage}{1.9cm}
  \begin{block}{}\centering
    \begin{tikzpicture}
      \tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
      \tikzstyle[level distance=.3cm]
      \tikzstyle[sibling distance=12cm]
      \tikzstyle{level 1}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=blue!15]}]
      \node{Total}[edge from parent fork down]
      child {node {X}
        }
      child {node {Y}
        };
    \end{tikzpicture}
  \end{block}
\end{minipage}
\end{textblock}

\begin{textblock}{3}(3.5,1.2)\alert{Bootstrap}\end{textblock}

##  Cross-temporal probabilistic forecast reconciliation
\vspace*{-0.2cm}

```{r}
#| label: bootres2
#| output: asis
#| dependson: bootres
boot_plot(c("red","blue","green","blue"))
```

\begin{textblock}{1.9}(12.9,5.3)
\begin{block}{}\centering
\textcolor[HTML]{7fb97f}{Year 1}\\
\textcolor[HTML]{7fbcff}{Year 2}\\
\textcolor[HTML]{e87f7f}{Year 3}\\
\textcolor[HTML]{7f7f7f}{Year 4}
\end{block}
\end{textblock}

\begin{textblock}{1.9}(12.9,1.6)
\begin{minipage}{1.9cm}
  \begin{block}{}\centering
    \begin{tikzpicture}
      \tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
      \tikzstyle[level distance=.3cm]
      \tikzstyle[sibling distance=12cm]
      \tikzstyle{level 1}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=blue!15]}]
      \node{Total}[edge from parent fork down]
      child {node {X}
        }
      child {node {Y}
        };
    \end{tikzpicture}
  \end{block}
\end{minipage}
\end{textblock}

\begin{textblock}{3}(3.5,1.2)\alert{Bootstrap}\end{textblock}

\only<2>{\begin{textblock}{8}(3.7,4)
\begin{alertblock}{}
The ``year'' can start in any quarter, giving overlapping blocks.
\end{alertblock}
\end{textblock}}

## Monthly Australian Tourism Demand

\begin{textblock}{6}(0.2,1.2)
\centering\fontsize{12}{13}\sf
\textbf{Geographical division}\\
\includegraphics[width = 5.5cm, trim= 0 0 180 0, clip=true]{subspace_projections_files/figure-beamer/ausmap-1.pdf}\\[-0.4cm]
\faTimes\\
\textbf{Purpose of travel}\\
{\fontsize{11}{12}\sf Holiday, Visiting friends \& relatives, Business, Other}
\end{textblock}

\begin{textblock}{10}(6.1,1)
\fontsize{11}{14}\sf\tabcolsep=0.12cm
\begin{itemize}
\item \textbf{Grouped ts}\newline (geographical divisions $\times$ purpose of travel)

\begin{tabular}{lccccc}
\toprule
  & \textbf{AUS} & \textbf{States} & \textbf{Zones$^\ast$} & \textbf{Regions} & \textbf{Tot}\\
  \midrule
  \textbf{geographical} & {\color{newblue}1} & {\color{newblue}7} & {\color{newblue}21} & {\color{newblue}76} & 105 \\
  \textbf{purpose} & {\color{newblue}4} & {\color{newblue}28} & {\color{newblue}84} & {\color{avocado}304} & 420\\
  \midrule
  \textbf{total} & 5 & 35 & 105 & 380 & \textbf{525}\\
  \bottomrule
\end{tabular}
\centerline{{\color{newblue}$n_a = 221$}, {\color{avocado}$n_b = 304$}, and $\textbf{n = 525}$}

\item \textbf{Temporal framework}, frequencies:\\[0.2cm]
\begin{multicols}{2}
  \begin{itemize}\tightlist
  \item Monthly
  \item Bi-Monthly
  \item Quarterly
  \end{itemize}
  \begin{itemize}\tightlist
  \item Four-Monthly
  \item Semi-Annual
  \item Annual
  \end{itemize}
\end{multicols}
\end{itemize}
\end{textblock}

## Monthly Australian Tourism Demand

* Monthly data: January 1998 to December 2016

* Time series cross-validation; initial training set 10 years.

* One-month increase in each training set

* For each training set, compute temporally aggregated series for $k \in \{1,2,3,4,6,12\}$, and produce forecasts up to $h_2=6$, $h_3=4$, $h_4=3$, $h_6=2$ and $h_{12}=1$ steps ahead.

* Automatic ETS forecasts on log-transformed data

## Monthly Australian Tourism Demand

\alert{Reconciliation approaches}\fontsize{13}{14}\sf

* Cross-temporal \textbf{\color{newblue}{bottom-up}} and \textbf{\color{newblue}{partly bottom-up}}

  \begin{tabular}{M{0.24\linewidth}|M{0.24\linewidth}|M{0.24\linewidth}}
  ct$(bu)$ & ct$(shr_{cs}, bu_{te})$ & ct$(wlsv_{te}, bu_{cs})$
  \end{tabular}

* Optimal forecast reconciliation with \textbf{\color{newblue}{one-step residuals}}

    \begin{tabular}{M{0.24\linewidth}|M{0.24\linewidth}|M{0.24\linewidth}|M{0.24\linewidth}}
    oct$(ols)$ & oct$(struc)$ & oct$(wlsv)$ & oct$(bdshr)$
    \end{tabular}

* Optimal forecast reconciliation with \textbf{\color{newblue}{multi-step residuals}}

    \begin{tabular}{M{0.24\linewidth}|M{0.24\linewidth}|M{0.24\linewidth}|M{0.24\linewidth}}
    oct$_h(hbshr)$ & oct$_h(bshr)$ & oct$_h(hshr)$ & oct$_h(shr)$
    \end{tabular}

## Monthly Australian tourism data -- CRPS skill scores
\fontsize{12}{13}\sf\vspace*{0.2cm}

\centerline{\textcolor{red}{Worse than benchmark}\qquad \textbf{Best}}

\centering
\footnotesize
\begin{tabular}[t]{lrr}
& \textbf{$\forall k \in \{12,6,4,3,2,1\}$} & \textbf{$k = 1$}\\
\midrule
base                      & \cellcolor{LightOrange!30} {1.000}  & \cellcolor{LightOrange!30} {1.000} \\
ct$(bu)$                  & \textcolor{red}{1.321}              & \textcolor{red}{1.077} \\
ct$(shr_{cs}, bu_{te})$   & \textcolor{red}{1.057}              & {0.976} \\
ct$(wlsv_{te}, bu_{cs})$  & \textcolor{red}{1.062}              & {0.976}\\
oct$(ols)$                & {0.989}                             & {0.982}\\
oct$(struc)$              & {0.982}                             & {0.970}\\
oct$(wlsv)$               & {0.987}                             & {0.952} \\
oct$(bdshr)$              & {0.975}                             & {\textbf{0.949}} \\
oct$_h(hbshr)$            & {0.989}                             & {0.982} \\
oct$_h(bshr)$             & {0.994}                             & {0.988}\\
oct$_h(hshr)$             & {\textbf{0.969}}                    & {0.953}\\
oct$_h(shr)$              & \textcolor{red}{1.007}              & \textcolor{red}{1.000}
\end{tabular}

# Improving multivariate forecasts

# Final comments

## Software
\fontsize{11}{12}\sf\vspace*{0.3cm}

\hspace*{-0.6cm}\begin{tabular}{llP{1.7cm}cP{1.7cm}c}
\toprule
Package                                                                      & Language  & Cross-sectional  & Temporal    & Cross-temporal  & Probabilistic\\
\midrule
\texttt{\href{https://pkg.earo.me/hts/}{hts}}
    & R         & \checkmark       &             &                 & \\
\texttt{\href{http://pkg.robjhyndman.com/thief/}{thief}}
    & R         &                  & \checkmark  &                 & \\
\texttt{\href{https://fable.tidyverts.org}{fable}}
    & R         & \checkmark       &             &                 & \checkmark\\
\texttt{\href{https://danigiro.github.io/FoReco/}{FoReco}}
    & R         & \checkmark       & \checkmark  & \checkmark      & \checkmark\\
\texttt{\href{https://angelpone.github.io/pyhts/}{pyhts}}
    & Python    & \checkmark       & \checkmark  &                 & \\
\texttt{\href{https://nixtla.github.io/hierarchicalforecast/}{hierarchicalforecast}}
    & Python    & \checkmark       &             &                 & \checkmark \\
\bottomrule
\end{tabular}

* `hts`, `thief`, and `FoReco` use `ts` objects
* `fable` uses `tsibble` objects
* `fable` has plans to implement temporal and cross-temporal reconciliation

## Thanks!

\placefig{0}{1.2}{trim = 10 45 0 0, clip=TRUE, width=10cm, height=2.5cm}{roman}
\placefig{2}{1.2}{trim = 0 40 0 0, clip=TRUE, width=10cm, height=2.5cm}{george}
\placefig{4}{1.2}{trim = 0 10 0 0, clip=TRUE, width=10cm, height=2.5cm}{hanlin}
\placefig{6}{1.2}{trim = 10 0 0 0, clip=TRUE, width=10cm, height=2.5cm}{earowang}
\placefig{8}{1.2}{trim = 0 15 0 0, clip=TRUE, width=10cm, height=2.5cm}{alanlee}
\placefig{10}{1.2}{trim = 30 0 0 0, clip=TRUE, width=10cm, height=2.5cm}{mitch}
\placefig{12}{1.2}{trim = 15 0 0 0, clip=TRUE, width=10cm, height=2.5cm}{shanika}
\placefig{14}{1.2}{trim = 40 0 0 0, clip=TRUE, width=10cm, height=2.5cm}{tas}

\placefig{0}{3.8}{trim = 30 10 30 0, clip=TRUE, width=10cm, height=2.5cm}{puwasala}
\placefig{2}{3.8}{trim = 0 10 0 0, clip=TRUE, width=10cm, height=2.5cm}{fotios}
\placefig{4}{3.8}{trim = 100 30 50 20, clip=TRUE, width=10cm, height=2.5cm}{nikos}
\placefig{6}{3.8}{trim = 50 30 0 0, clip=TRUE, width=10cm, height=2.5cm}{souhaib}
\placefig{8}{3.8}{trim = 110 40 50 0, clip=TRUE, width=10cm, height=2.5cm}{james}
\placefig{10}{3.8}{trim = 40 40 0 0, clip=TRUE, width=10cm, height=2.5cm}{mahdi}
\placefig{12}{3.8}{trim = 50 50 0 0, clip=TRUE, width=10cm, height=2.5cm}{christoph}
\placefig{14}{3.8}{trim = 50 50 0 20, clip=TRUE, width=10cm, height=2.5cm}{fin}

\placefig{0}{6.4}{trim = 10 0 0 0, clip=TRUE, width=10cm, height=2.5cm}{berwin}
\placefig{2}{6.4}{trim = 10 20 0 0, clip=TRUE, width=10cm, height=2.5cm}{galit}
\placefig{4}{6.4}{trim = 10 0 0 0, clip=TRUE, width=10cm, height=2.5cm}{mahsa}
\placefig{6}{6.4}{trim = 30 0 0 0, clip=TRUE, width=10cm, height=2.5cm}{evan}
\placefig{8}{6.4}{trim = 5 25 0 0, clip=TRUE, width=10cm, height=2.5cm}{bahman}
\placefig{10}{6.4}{trim = 0 40 0 0, clip=TRUE, width=10cm, height=2.5cm}{pablo}
\placefig{12}{6.4}{trim = 0 40 0 0, clip=TRUE, width=10cm, height=2.5cm}{danielegiro}
\placefig{14}{6.4}{trim = 0 0 0 30, clip=TRUE, width=10cm, height=2.5cm}{tommy}

## More information
\fontsize{18}{20}\sf

\href{https://robjhyndman.com}{\faicon{home} robjhyndman.com}

\href{https://aus.social/@robjhyndman}{\includegraphics[width=0.5cm]{figs/mastodon}\, aus.social/@robjhyndman}

\href{https://github.com/robjhyndman}{\faicon{github}  @robjhyndman}

\href{mailto:rob.hyndman@monash.edu}{\faicon{envelope}  rob.hyndman@monash.edu}

\nocite{Di_FonGir2022a,temporal-hierarchies,ctprob}
