---
title: Improving forecasts via subspace projections
author: Rob J Hyndman
date: 19 September 2024
toc: false
format:
  presentation-beamer:
    pdf-engine: pdflatex
    knitr:
      opts_chunk:
        dev: "CairoPDF"
    template-partials:
      - before-title.tex
    include-in-header: header.tex
    keep-tex: true
    cite-method: biblatex
    fig-width: 9
    fig-height: 4
biblatexoptions: natbib,style=authoryear
bibliography: hts.bib
highlight-style: tango
execute:
  echo: false
  message: false
  warning: false
  cache: false
abstract: Univariate, multivariate, and hierarchical forecasts can all be improved using projections onto linear subspaces, regardless of what forecasting method is used. I will show some theoretical guarantees of this statement, and demonstrate using empirical applications how linear projections can lead to (sometimes dramatic) improvements in forecast accuracy.
---

```{r}
#| label: load-packages
source("setup.R")
library(fpp3)
library(targets)

tar_load(c(tourism, tourism_agg, fit, fc))
```

## Sri Lankan PhD students and post-docs

\placefig{0}{1.2}{height = 3.2cm}{arivalzahan}
\placefig{3.2}{1.2}{height = 3.2cm}{thilaksha}
\placefig{6.4}{1.2}{height = 3.2cm}{shanika}
\placefig{9.6}{1.2}{height = 3.2cm}{dilini}
\placefig{12.8}{1.2}{height = 3.2cm}{thiyanga}
\placefig{0}{4.5}{height = 3.2cm}{sevvandi}
\placefig{3.2}{4.5}{height = 3.2cm}{puwasala}
\placefig{6.4}{4.5}{height = 3.2cm}{nuwani}
\placefig{9.6}{4.5}{height = 3.2cm}{lakshan}

## Outline

\vspace*{0.7cm}
\tableofcontents[hideallsubsections]

# Improving hierarchical forecasts

## Australian tourism regions

\placefig{0.3}{1.2}{height=7.7cm}{figs/ausmap.png}

\only<2>{\begin{textblock}{6.5}(9.2,-0.15)
\begin{block}{}\fontsize{12}{13}\sf
  \begin{itemize}\itemsep=0cm\parskip=0cm
    \item Monthly data on visitor nights: 1998 -- 2016
    \item 7 states
    \item 27 zones
    \item 76 regions
  \end{itemize}
\end{block}
\end{textblock}}

## Australian tourism data

\only<1>{\placefig{0.1}{1.4}{width=15.8cm, height=7.5cm, keepaspectratio}{tourism1}}
\only<2>{\placefig{0.1}{1.4}{width=15.8cm, height=7.5cm, keepaspectratio}{tourism2}}
\only<3>{\placefig{0.1}{1.4}{width=15.8cm, height=7.5cm, keepaspectratio}{tourism3}}
\only<4>{\placefig{0.1}{1.4}{width=15.8cm, height=7.5cm, keepaspectratio}{tourism4}}

## Australian tourism data

\placefig{0.2}{1.3}{width = 6.8cm, trim = 0 0 700 0, clip = true}{ausmap.png}

\begin{textblock}{10}(8.1,1.4)
\fontsize{11}{12}\sf
\begin{tabular}{lr}
\toprule
\bf Aggregation level & \bf \# series \\
\midrule
National & 1 \\
State & 7 \\
Zone & 27 \\
Region & 76 \\
\midrule
\bf Total & \bf 111 \\
\bottomrule
\end{tabular}
\end{textblock}

\begin{textblock}{9.2}(6.5,5.5)
\begin{alertblock}{}\fontsize{12}{15}\sf
\begin{itemize}
\item Need forecasts at all levels of aggregation.
\item Compute \textbf{base forecasts} using univariate models. These will not add up.
\item Adjust base forecasts to ensure they are ``coherent'' giving \textbf{reconciled forecasts}.
\end{itemize}
\end{alertblock}
\end{textblock}

## Notation

\begin{textblock}{8.8}(0.2,1.5)
Almost all collections of time series with linear constraints can be written \rlap{as}
\centerline{\colorbox[RGB]{210,210,210}{$\bY_{t}=\color{summing}\bS\color{bottom}\bm{b}_{t}$}}
\vspace*{-0.9cm}\begin{itemize}\parskip=0cm\itemsep=0cm
\item $\by_t=$ vector of all series at time $t$
\item $y_{\text{Total},t}=$ aggregate of all series at time $t$.
\item $y_{X,t}=$ value of series $X$ at time $t$.
\item $\color{bottom}{\bm{b}_t}=$ vector of most disaggregated series at time $t$
\item $\color{summing}{\bS}=$ ``summing matrix'' containing the linear constraints.
\end{itemize}
\end{textblock}

\begin{textblock}{5.7}(11.4,0.1)
\begin{minipage}{4cm}
\begin{block}{}\centering
\begin{tikzpicture}
\tikzstyle{every node}=[ellipse,draw,fill=red!15,inner sep=2pt]
\tikzstyle[level distance=.3cm]
\tikzstyle[sibling distance=12cm]
\tikzstyle{level 1}=[sibling distance=10mm,font=\small,set style={{every node}+=[fill=blue!15]}]
\node{Total}[edge from parent fork down]
 child {node {A}
 }
 child {node {B}
 };
\end{tikzpicture}
\end{block}
\end{minipage}
\end{textblock}

\begin{textblock}{5.7}(9.4,2.8)\fontsize{15}{17}\sf
\begin{align*}
\bY_{t}&= \begin{pmatrix}
  y_{\text{Total},t}\\
  y_{A,t}\\
  y_{B,t}
  \end{pmatrix}  \\
  &= {\color{summing}\underbrace{\begin{pmatrix}
                1 & 1  \\
                1 & 0  \\
                0 & 1
                \end{pmatrix}}_{\bm{S}}}
     {\color{bottom}\underbrace{\begin{pmatrix}
       y_{A,t}\\y_{B,t}
       \end{pmatrix}}_{\bm{b}_t}}
\end{align*}
\end{textblock}

## Projections onto the coherent subspace

\placefig{0}{1.2}{width=7.4cm}{3D_hierarchy}

\begin{textblock}{3}(2,7.5)\fontsize{13}{13}\sf
\begin{block}{}
\centerline{$y_{\text{Total}} = y_A + y_B$}
\end{block}
\end{textblock}

\only<2->{\begin{textblock}{7.5}(8.3,1)\fontsize{13}{13}\sf
\begin{alertblock}{Base forecasts}
Let $\base$ be vector of \emph{incoherent} initial $h$-step forecasts.
\end{alertblock}
\end{textblock}}

\only<3->{\begin{textblock}{7.5}(8.3,3.1)\fontsize{13}{13}\sf
\begin{alertblock}{Reconciled forecasts}
Let $\bm{M}$ be a projection matrix. $\reconciled = \bm{M} \base$ ``reconciles'' $\base$.
\end{alertblock}
\end{textblock}}

\only<4>{\begin{textblock}{7.8}(8,5.4)\fontsize{13}{15}\sf
\begin{itemize}\tightlist
\item $\bm{S}$ forms a basis set for $\mathfrak{s}$
\item All projections are of the form $\bm{M} = \bS(\bS'\bm{\Psi}\bS)^{-1}\bS'\bm{\Psi}$ where $\bm{\Psi}$ is a positive definite matrix.
\item How to choose the best $\bm{\Psi}$?
\end{itemize}
\end{textblock}}

## Minimum trace reconciliation

\begin{textblock}{6.4}(9,-0.1)\begin{block}{}
Wickramasuriya et al (2019)
\end{block}\end{textblock}

\begin{block}{}
\centerline{$\displaystyle\reconciled = \bm{M}\base$}
\end{block}
\centerline{\hspace*{1.4cm}\textcolor{reconciled}{Reconciled forecasts}\hfill\textcolor{base}{Base forecasts}\hspace*{2.6cm}}
\pause

* Base forecast covariance:\newline $\bm{W}_h = \var[\by_{T+h} - \base \mid \by_1,\dots,\by_T]$
* Reconciled forecast covariance:\newline $\bm{V}_h = \var[\by_{T+h} - \reconciled \mid \by_1,\dots,\by_T])  = \bm{M}\bm{W}_h\bm{M}'$

\pause
\begin{alertblock}{Minimum trace (MinT) reconciliation}
If $\bm{M}$ is a projection, then trace of $\bm{V}_h$ is minimized when $\bm{\Psi} = \bm{W}_h$
\centerline{$\bm{M} = \bS(\bS'\bm{W}_h^{-1}\bS)^{-1}\bS'\bm{W}_h^{-1}$}
\end{alertblock}

## Linear projections

\begin{textblock}{7}(7,-0.2)
\begin{block}{}
\centerline{$\reconciled = \bS(\bS'\bm{W}_h^{-1}\bS)^{-1}\bS'\bm{W}_h^{-1}\base$}
\end{block}
\end{textblock}

\begin{textblock}{9.4}(1,1.15)
\begin{alertblock}{Reconciliation method \hspace*{0.5cm} $\bm{W}_h$}
\begin{tabular}{ll}
  OLS             & $\bm{I}$ \\
  WLS(var)        & $\text{diag}(\bm{W}_1)$ \\
  WLS(struct)     & $\text{diag}(\bS\bm{1})$ \\
  MinT(sample)    & $\hat{\bm{W}}_{\text{sam}}$  \\
  MinT(shrink)\hspace*{2.5cm}    & $\hat{\bm{W}}_{\text{shr}}$  
\end{tabular}
\end{alertblock}
\end{textblock}
\begin{textblock}{15}(.8,5.8)\fontsize{14}{16}\sf
\begin{itemize}\parskip=0cm
\item All approximate MinT by assuming $\bm{W}_h = k_h \bm{W}_1$.
\item $\hat{\bm{W}}_{\text{sam}}$ is sample estimate of the residual covariance matrix
\item $\hat{\bm{W}}_{\text{shr}}$ is shrinkage estimator $\tau \text{diag}(\hat{\bm{W}}_{\text{sam}})+(1-\tau)\hat{\bm{W}}_{\text{sam}}$\\ where $\tau$ selected optimally.
\end{itemize}
\end{textblock}


## Example: Australian tourism
\fontsize{10}{11}\sf

```{r tourismdata00, echo=TRUE}
tourism
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```r
tourism_agg <- tourism |>
  aggregate_key(state / zone / region, visitors = sum(visitors))
```

```{r tourismagg2, echo=FALSE}
#| label: tourismagg
tourism_agg
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```r
fit <- tourism_agg |>
  filter(year(month) <= 2015) |>
  model(ets = ETS(visitors))
```

```{r tourismmodels1, echo=FALSE}
#| label: tourismmodels
fit
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```r
fc <- fit |>
  reconcile(
    ols = min_trace(ets, method = "ols"),
    mint_s = min_trace(ets, method = "mint_shrink"),
  ) |>
  forecast(h = "2 years")
```

```{r fctourism1, echo=FALSE}
#| label: fctourism
fc
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```r
fc |>
  filter(is_aggregated(state)) |>
  autoplot(filter(tourism_agg, year(month) > 2012), level = NULL)
```

```{r}
#| label: fctourism2
#| fig-width: 12
#| fig-height: 4.5
tar_read(fctourism2)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```r
fc |>
  filter(state == "NSW" & is_aggregated(zone)) |>
  autoplot(filter(tourism_agg, year(month) > 2012), level = NULL)
```

```{r}
#| label: fctourism3
#| fig-width: 12
#| fig-height: 4.5
tar_read(fctourism3)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```r
fc |>
  filter(region == "Melbourne") |>
  autoplot(filter(tourism_agg, year(month) > 2012), level = NULL)
```

```{r}
#| label: fctourism4
#| fig-width: 12
#| fig-height: 4.5
tar_read(fctourism4)
```


## Example: Australian tourism
\fontsize{10}{11}\sf

```r
fc |>
  filter(region == "Snowy Mountains") |>
  autoplot(filter(tourism_agg, year(month) > 2012), level = NULL)
```

```{r}
#| label: fctourism5
#| fig-width: 12
#| fig-height: 4.5
tar_read(fctourism5)
```


## Example: Australian tourism
\fontsize{10}{11}\sf

```r
fc |>
  filter(region == "Barossa") |>
  autoplot(filter(tourism_agg, year(month) > 2012), level = NULL)
```

```{r}
#| label: fctourism6
#| fig-width: 12
#| fig-height: 4.5
tar_read(fctourism6)
```

## Performance evaluation

\vspace*{0.2cm}\begin{block}{}
\centerline{$\text{RMSSE} = \sqrt{\text{mean}(q_{j}^2)}$}
$$
  q^2_{j} = \frac{ e^2_{j}}
 {\displaystyle\frac{1}{T-m}\sum_{t=m+1}^T (y_{t}-y_{t-m})^2}
$$
\end{block}

* $y_t=$ observation for period $t$
* $e_{j}=$ forecast error for forecast horizon $j$
* $T=$ size of training set
* $m = 12$

## Example: Australian tourism
\fontsize{10}{11}\sf

```r
accuracy(fc, tourism_agg, measures = list(rmsse = RMSSE))
```

```{r}
#| label: fcaccuracy
tar_read(fcaccuracy)
```

## Example: Australian tourism
\fontsize{10}{11}\sf

```r
accuracy(fc, tourism_agg, measures = list(rmsse = RMSSE)) |>
  summarise(rmsse = sqrt(mean(rmsse^2)), .by = .model)
```

```{r}
#| label: fcaccuracy2
tar_read(fcaccuracy2)
```

\only<2>{\begin{textblock}{7}(7,3)
\begin{block}{}\fontsize{13}{14}\sf
\begin{itemize}\tightlist
\item Overall, both reconciliation methods are more accurate than the base ETS forecasts.
\end{itemize}
\end{block}
\end{textblock}}

## Example: Australian tourism
\fontsize{8}{7}\sf


```r
accuracy(fc, tourism_agg, measures = list(rmsse = RMSSE)) |>
  summarise(rmsse = sqrt(mean(rmsse^2)), .by = c(.model, level)) |>
```

```{r fcaccuracy3, echo=FALSE}
tar_read(fcaccuracy3)
```

\only<2>{\begin{textblock}{7}(7,3)
\begin{block}{}\fontsize{13}{14}\sf
\begin{itemize}\tightlist
\item Reconciliation is best for all levels except national.
\item Improvements due to reconciliation are greater at lower levels.
\end{itemize}
\end{block}
\end{textblock}}

## Mean square error bounds

\begin{textblock}{6.4}(9,-0.1)\begin{block}{}
\citet{htsgeometry}
\end{block}\end{textblock}

\vspace*{0.2cm}\begin{alertblock}{Distance reducing property}
Let $\|\bm{u}\|_{\bm{\Psi}} = \bm{u}'\bm{\Psi}\bm{u}$. Then
  \centerline{$\|\bm{y}_{t+h}-\reconciled\|_{\bm{\Psi}}\le\|\bm{y}_{t+h}-\base\|_{\bm{\Psi}}$}
\end{alertblock}

 * $\bm{\Psi}$-projection is guaranteed to improve forecast accuracy over base forecasts *using this distance measure*.
 * Distance reduction holds for any realisation and any forecast.
 * OLS reconciliation minimizes Euclidean distance.

## Mean square error bounds

\begin{textblock}{6.4}(9,-0.1)\begin{block}{}
\citet{wickramasuriya2021properties}
\end{block}\end{textblock}

\begin{block}{}\vspace*{-0.6cm}
\begin{align*}
\|\bm{y}_{t+h} - \reconciled\|_2^2
 &= \|\bm{M}(\bm{y}_{t+h} - \base)\|_2^2 \\
 &\le \|\bm{M}\|_2^2 \|\bm{y}_{t+h} - \base\|_2^2 \\
 & = \sigma_{\text{max}}^2\|\bm{y}_{t+h} - \base\|_2^2
\end{align*}
\end{block}

 * $\sigma_{\text{max}}$ is the largest eigenvalue of $\bm{M}$
 * $\sigma_{\text{max}}\ge1$ as $\bm{M}$ is a projection matrix.
 * Every projection reconciliation is better than base forecasts using Euclidean distance.

## Mean square error bounds

\begin{textblock}{6.4}(9,-0.1)\begin{block}{}
\citet{wickramasuriya2021properties}
\end{block}\end{textblock}
\vspace*{0.2cm}\begin{block}{}\vspace*{-0.6cm}
\begin{align*}
    & \text{tr}\Big(\E[\bm{y}_{t+h} - \reconciled^{\color{reconciled}\text{MinT}}]'[\bm{y}_{t+h} - \reconciled^{\color{reconciled}\text{MinT}}]\Big) \\
\le~ & \text{tr}\Big(\E[\bm{y}_{t+h} - \reconciled^{\color{reconciled}\text{OLS}}]'[\bm{y}_{t+h} - \reconciled^{\color{reconciled}\text{OLS}}]\Big) \\
\le~ & \text{tr}\Big(\E[\bm{y}_{t+h} - \base]'[\bm{y}_{t+h} - \base]\Big)
\end{align*}
\end{block}

Using sums of variances:

* MinT reconciliation is better than OLS reconciliation
* OLS reconciliation is better than base forecasts

# Improving univariate forecasts

## Temporal reconciliation: quarterly data

\begin{textblock}{7}(1,1.9)
\begin{tikzpicture}
  \tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15,font=\small]
  \tikzstyle[level distance=.1cm]
  \tikzstyle[sibling distance=3cm]
  \tikzstyle{level 1}=[sibling distance=18mm,set style={{every node}+=[fill=yellow]}]
  \node{Annual}[edge from parent fork down]
  child {node {Q$_1$}}
  child {node {Q$_2$}}
  child {node {Q$_3$}}
  child {node {Q$_4$}};
\end{tikzpicture}
\end{textblock}

\only<3>{\begin{textblock}{8.5}(8.3,1.1)\fontsize{14}{17}\sf
$$\bm{y}_\tau =
  \begin{bmatrix}
    x_{\tau}^{[4]}  \\[0.2cm]
    x_{\tau,1}^{[1]} \\[0.2cm]
    x_{\tau,2}^{[1]} \\[0.2cm]
    x_{\tau,3}^{[1]} \\[0.2cm]
    x_{\tau,4}^{[1]}
  \end{bmatrix}
  \qquad
  \bm{S} = \begin{bmatrix}
    1 & 1 & 1 & 1 \\[0.2cm]
    1 & 0 & 0 & 0 \\[0.2cm]
    0 & 1 & 0 & 0 \\[0.2cm]
    0 & 0 & 1 & 0 \\[0.2cm]
    0 & 0 & 0 & 1
  \end{bmatrix}
$$
\end{textblock}}

\only<2->{\begin{textblock}{8}(.3,5.7)
  \begin{alertblock}{}
    \begin{itemize}
      \item[\color{white}\ding{229}] Forecast series at each available frequency.
      \item[\color{white}\ding{229}] Optimally combine forecasts within the same year.
    \end{itemize}
  \end{alertblock}
\end{textblock}}

\only<3->{\begin{textblock}{5}(10.3,7.9)
\fontsize{11}{12}\sf
$\tau=$ index of largest temporal aggregation level.
\end{textblock}}


## Temporal reconciliation: quarterly data

\begin{textblock}{7}(0.1,1.9)
\begin{tikzpicture}
  \tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15,font=\small]
  \tikzstyle[level distance=.1cm]
  \tikzstyle[sibling distance=3cm]
  \tikzstyle{level 1}=[sibling distance=42mm,set style={{every node}+=[fill=blue!15]}]
  \tikzstyle{level 2}=[sibling distance=18mm,set style={{every node}+=[fill=yellow]}]
  \node{Annual}[edge from parent fork down]
  child {node {Semi-Annual$_1$}
      child {node {Q$_1$}}
      child {node {Q$_2$}}
    }
  child {node {Semi-Annual$_2$}
      child {node {Q$_3$}}
      child {node {Q$_4$}}
    };
\end{tikzpicture}
\end{textblock}

\begin{textblock}{8.5}(8.3,1.1)\fontsize{14}{17}\sf
$$\bm{y}_\tau =
  \begin{bmatrix}
    x_{\tau}^{[4]}  \\[0.2cm]
    x_{\tau,1}^{[2]} \\[0.2cm]
    x_{\tau,2}^{[2]} \\[0.2cm]
    x_{\tau,1}^{[1]} \\[0.2cm]
    x_{\tau,2}^{[1]} \\[0.2cm]
    x_{\tau,3}^{[1]} \\[0.2cm]
    x_{\tau,4}^{[1]}
  \end{bmatrix}
  \qquad
  \bm{S} = \begin{bmatrix}
    1 & 1 & 1 & 1 \\[0.2cm]
    1 & 1 & 0 & 0 \\[0.2cm]
    0 & 0 & 1 & 1 \\[0.2cm]
    1 & 0 & 0 & 0 \\[0.2cm]
    0 & 1 & 0 & 0 \\[0.2cm]
    0 & 0 & 1 & 0 \\[0.2cm]
    0 & 0 & 0 & 1
  \end{bmatrix}
$$
\end{textblock}

\begin{textblock}{8}(.3,5.7)
  \begin{alertblock}{}
    \begin{itemize}
      \item[\color{white}\ding{229}] Forecast series at each available frequency.
      \item[\color{white}\ding{229}] Optimally combine forecasts within the same year.
    \end{itemize}
  \end{alertblock}
\end{textblock}

\begin{textblock}{5}(10.3,7.9)
\fontsize{11}{12}\sf
$\tau=$ index of largest temporal aggregation level.
\end{textblock}

## Temporal reconciliation: monthly data

\only<1>{\begin{tikzpicture}
  \tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15,font=\small]
  \tikzstyle[level distance=.1cm]
  \tikzstyle[sibling distance=7cm]
  \tikzstyle{level 1}=[sibling distance=72mm,set style={{every node}+=[fill=blue!15]}]
  \tikzstyle{level 2}=[sibling distance=36mm,set style={{every node}+=[fill=yellow]}]
  \tikzstyle{level 3}=[sibling distance=12mm,font=\scriptsize,set style={{every node}+=[fill=green]}]
  \node{Annual}[edge from parent fork down]
  child {node {Semi-Annual$_1$}
      child {node {Q$_1$}
          child {node {\scriptsize M$_1$}}
          child {node {\scriptsize M$_2$}}
          child {node {\scriptsize M$_3$}}
        }
      child {node {Q$_2$}
          child {node {\scriptsize M$_4$}}
          child {node {\scriptsize M$_5$}}
          child {node {\scriptsize M$_6$}}
        }
    }
  child {node {Semi-Annual$_2$}
      child {node {Q$_3$}
          child {node {\scriptsize M$_7$}}
          child {node {\scriptsize M$_8$}}
          child {node {\scriptsize M$_9$}}
        }
      child {node {Q$_4$}
          child {node {\scriptsize M$_{10}$}}
          child {node {\scriptsize M$_{11}$}}
          child {node {\scriptsize M$_{12}$}}
        }
    };
\end{tikzpicture}}
\only<2->{\begin{tikzpicture}
  \tikzstyle{every node}=[ellipse,draw,inner sep=0.2pt,fill=red!15,font=\small]
  \tikzstyle[level distance=.1cm]
  \tikzstyle[sibling distance=7cm]
  \tikzstyle{level 1}=[sibling distance=48mm,set style={{every node}+=[fill=blue!15]}]
  \tikzstyle{level 2}=[sibling distance=24mm,set style={{every node}+=[fill=yellow]}]
  \tikzstyle{level 3}=[sibling distance=12mm,set style={{every node}+=[fill=green]}]
  \node{Annual}[edge from parent fork down]
  child {node {FourM$_1$}
      child {node {BiM$_1$}
          child {node {\scriptsize M$_1$}}
          child {node {\scriptsize M$_2$}}
        }
      child {node {BiM$_2$}
          child {node {\scriptsize M$_3$}}
          child {node {\scriptsize M$_4$}}
        }
    }
  child {node {FourM$_2$}
      child {node {BiM$_3$}
          child {node {\scriptsize M$_5$}}
          child {node {\scriptsize M$_6$}}
        }
      child {node {BiM$_4$}
          child {node {\scriptsize M$_7$}}
          child {node {\scriptsize M$_8$}}
        }
    }
  child {node {FourM$_3$}
      child {node {BiM$_5$}
          child {node {\scriptsize M$_9$}}
          child {node {\scriptsize M$_{10}$}}
        }
      child {node {BiM$_6$}
          child {node {\scriptsize M$_{11}$}}
          child {node {\scriptsize M$_{12}$}}
        }
    };
\end{tikzpicture}}\pause

\begin{textblock}{14}(1,6.7)
  \begin{alertblock}{}
    \begin{itemize}
      \item[\color{white}\ding{229}] Forecast series at each available frequency.
      \item[\color{white}\ding{229}] Optimally combine forecasts within the same year.
    \end{itemize}
  \end{alertblock}
\end{textblock}

## Temporal reconciliation: monthly data
\fontsize{11}{11}\sf

$$
  \bm{y}_\tau=\begin{bmatrix}
    x_{\tau}^{[12]}     \\[0.2cm]
    \bm{x}_{\tau}^{[6]} \\[0.2cm]
    \bm{x}_{\tau}^{[4]} \\[0.2cm]
    \bm{x}_{\tau}^{[3]} \\[0.2cm]
    \bm{x}_\tau^{[2]}   \\[0.2cm]
    \bm{x}_\tau^{[1]}
  \end{bmatrix}
  \qquad
  \bm{S} = \begin{bmatrix}
    1                & 1 & 1 & 1 & 1~~~1~~~1~~~1 & 1 & 1 & 1 & 1 \\
    1                & 1 & 1 & 1 & 1~~~1~~~0~~~0 & 0 & 0 & 0 & 0 \\
    0                & 0 & 0 & 0 & 0~~~0~~~1~~~1 & 1 & 1 & 1 & 1 \\
    1                & 1 & 1 & 1 & 0~~~0~~~0~~~0 & 0 & 0 & 0 & 0 \\
    0                & 0 & 0 & 0 & 1~~~1~~~1~~~1 & 0 & 0 & 0 & 0 \\
    0                & 0 & 0 & 0 & 0~~~0~~~0~~~0 & 1 & 1 & 1 & 1 \\
    1                & 1 & 1 & 0 & 0~~~0~~~0~~~0 & 0 & 0 & 0 & 0 \\
                     &   &   &   & \vdots        &   &   &   &   \\
    0                & 0 & 0 & 0 & 0~~~0~~~0~~~0 & 0 & 1 & 1 & 1 \\
    1                & 1 & 0 & 0 & 0~~~0~~~0~~~0 & 0 & 0 & 0 & 0 \\
                     &   &   &   & \vdots        &   &   &   &   \\
    0                & 0 & 0 & 0 & 0~~~0~~~0~~~0 & 0 & 0 & 1 & 1 \\[0.2cm]
    \phantom{\vdots} &   &   &   & \bm{I}_{12}   &   &   &   &
  \end{bmatrix}
$$

## Temporal reconciliation
\fontsize{14}{15}\sf

For a time series  $y_1,\dots,y_T$, observed at frequency $m$:
\begin{alertblock}{}\vspace*{-0.1cm}
$$
  x_j^{[k]} = \sum_{t = (j-1)k+1}^{jk} y_t\qquad \text{for $j = 1,\dots,\lfloor T/k\rfloor$}
$$
\end{alertblock}
* $k \in \mathcal{K} = \{k_1,\dots,k_p\}$ denote the $p$ factors of $m$ in ascending order, where $k_1=1$ and $k_p=m$
* $x_j^{[1]} = y_t$
* A single unique hierarchy is only possible when there are no coprime pairs in $\mathcal{K}$.
* $M_k=m/k$ is seasonal period of aggregated series.

## Temporal reconciliation
\fontsize{14}{15}\sf\vspace*{-0.5cm}
$$\bm{x}_\tau = \bm{S} \bm{x}_\tau^{[1]}, \qquad \bm{S} = \begin{bmatrix}\bm{A}\\\bm{I}\end{bmatrix}$$
where
$$
\bm{x}_\tau = \begin{bmatrix*}[l]
    {x_\tau^{[k_p]}}\\
    {\bm{x}_\tau^{[k_{p-1}]}}\\
    \quad\vdots\\
    {\bm{x}_\tau^{[k_1]}}\\
  \end{bmatrix*}\qquad
  \bm{x}_\tau^{[k]} = \begin{bmatrix*}[l]
    x_{M_k(\tau-1)+1}^{[k]}\\
    x_{M_k(\tau-1)+2}^{[k]}\\
    \quad\vdots\\
    x_{M_k\tau}^{[k]}
  \end{bmatrix*}\qquad
\bm{A} =
  \begin{bmatrix}
    \bm{1}'_m                                    \\
    \bm{I}_{m/k_{p-1}} \otimes \bm{1}'_{k_{p-1}} \\
    \vdots                                       \\
    \bm{I}_{m/k_{2}} \otimes \bm{1}'_{k_{2}}     \\
  \end{bmatrix}
$$
$\tau$ is time index for most aggregated series,

$k\in \mathcal{K} = \{k_1,\dots,k_p\}$,\quad $k_1=1$,\quad $k_p=m$,\quad $\tau=1,\dots,T/m$.



## Example: Accident & emergency services demand

Weekly A\&E demand data: 7 November 2010 to 7 June 2015.\fontsize{11}{11.5}\sf

\begin{tabular}{l}
  \toprule
  Type 1 Departments --- Major A\&E                               \\
  Type 2 Departments --- Single Specialty                         \\
  Type 3 Departments --- Other A\&E/Minor Injury Unit             \\
  Total Attendances                                               \\
  Type 1 Departments --- Major A\&E $>$ 2 hours                   \\
  Type 2 Departments --- Single Specialty $>$ 2 hours             \\
  Type 3 Departments --- Other A\&E/Minor Injury Unit $>$ 2 hours \\
  Total Attendances $>$ 2 hours                                   \\
  Emergency Admissions via Type 1 A\&E                            \\
  Total Emergency Admissions via A\&E                             \\
  Other Emergency Admissions (i.e not via A\&E)                   \\
  Total Emergency Admissions                                      \\
  Number of patients spending $>$ 2 hours from decision to admit to admission
\end{tabular}

## Example: Accident & emergency services demand

\placefig{1}{2.2}{height=6.5cm, width=14cm}{AEexample.pdf}

\begin{textblock}{10}(3,1)\fontsize{12}{12}\sf
  \begin{block}{}\centering
    Total emergency admissions via A\&E
  \end{block}
\end{textblock}

## Example: Accident & emergency services demand

Test set: last 52 weeks

\alert{MASE comparison} (ARIMA models)

```{r}
tar_read(aetable)
```

## Temporal reconciliation: M3 monthly series

* Apply temporal reconciliation to all 1428 monthly series from M3 competition
* Forecast horizon $h=18$ months
* ETS and ARIMA models
* Measure percentage difference to base forecasts
* Reconciliation methods:
  * WLS$_H$ (diagonal)
  * WLS$_V$ (diagonal with common variances for same frequency)
  * WLS$_S$ (diagonal/structural)

## Temporal reconciliation: M3 monthly series

\alert{Improvement in MASE relative to base forecasts}

\fontsize{11}{11}\sf
\tabcolsep=0.15cm
\centering
```{r}
options(knitr.kable.NA = "")
tar_read(tabm3)
```

# Improving multivariate forecasts

## Forecast Linear Augmented Projection (FLAP)

* We want to forecast multivariate series $\bm{y}_t$.
* Linear combinations of $\bm{y}_t$ may have better signal-noise ratio
* Construct many linear combinations $\bm{c}_{t} = \bm{\Phi}\bm{y}_t$ (e.g., principal components or random combinations)
* Produce univariate forecasts of all series $\base$ and all linear combinations $\textcolor{base}{\hat{\bm{c}}_{t+h|t}}$.
* Reconcile forecasts so they are coherent ($\textcolor{reconciled}{\tilde{\bm{c}}_{t+h|t}} = \bm{\Phi}\reconciled$)

\pause\vspace*{-0.4cm}

$$
\bm{z}_t = \begin{bmatrix} \bm{y}_t\\ \bm{c}_t \end{bmatrix}
\qquad \textcolor{reconciled}{\tilde{\bm{z}}_{t+h}} = \bm{M} \basez
$$
where $\bm{M}$ is a projection matrix onto the coherent subspace.

## Forecast error variance reduction

If we know the covariance matrix $\bm{W}_h = \Var(\bm{z}_{t+h} - \basez)$, then 

1. The forecast error variance is reduced:

   * $\Var(\bm{y}_{t+h} - \base) -\Var(\bm{y}_{t+h} - \reconciled)$ is positive semi-definite.
  
2. The forecast error variance of each series monotonically decreases with increasing number of components.

\pause

In practice, we need to:

* Estimate $\bm{W}_h$ (use the MinT shrinkage estimator).
* Construct the components, $\bm{\Phi}$.

## Construction of $\bm{\Phi}$

### Principal component analysis (PCA)

Finding the weights matrix $\bm{\Phi}$ so that the resulting components \alert{\textbf{maximise variance}}

### Simulation

Generating values of $\bm{\Phi}$ from a random distribution and normalising them to unit vectors

* Normal distribution
* Uniform distribution
* Orthonormal matrix

## Monthly Australian regional tourism

* Monthly Australian tourism data set aggregated by region giving 77 series, from Jan 1998 to Dec 2019.

* Use expanding window time series cross-validation with $T=84$ observations in first training set, and forecast horizons $h=1,2,\dots,12$.

* Fit univariate ETS models to each series.

## Monthly Australian regional tourism

```{r}
#| label: series
#| fig-height: 4.8
tar_read(series)
```

## Monthly Australian regional tourism

```{r}
#| label: components
#| fig-height: 4.8
tar_read(components_plot)
```

## Monthly Australian regional tourism

```{r visnights}
tar_read(visnights_mse)
```

## FRED-MD

* Monthly data of macroeconomic variables (McCracken and Ng, 2016).

* Data from Jan 1959 -- Sep 2023. 777 observations on 122 series.

* Same cleaning process as per McCracken and Ng (2016).

* All series scaled to have mean 0 and variance 1.

* Expanding time series cross-validation with initial size of 25 years and forecast horizon 12 months.

## FRED-MD (ARIMA)

```{r}
#| label: fred-md-arima
tar_read(plot_fred_mse_arima)
```

## FRED-MD (DFM)

```{r}
#| label: fred-md-dfm
tar_read(plot_fred_mse_dfm)
```

## Simulation

* Data generating process: VAR($3$) with 70 variables

* Sample size: $T=400$

* Number of repeated samples: $220$

* Base models:

  * automatic ARIMA (based on AICc)
  * DFM (structure chosen using BIC, different model for each horizon)

## Simulation

```{r simulation}
tar_read(plot_sim_mse)
```

## Future research directions

* Investigate why PCA performs better than random weights

* Find other components that are better than PCA

* Find optimal components by minimising forecast error variance with respect to $\bm{\Phi}$

* Use forecast projection and forecast reconciliation together


# Final comments

## Software
\fontsize{10}{12}\sf\vspace*{0.3cm}\tabcolsep=0.12cm

\hspace*{-0.3cm}\begin{tabular}{@{}llP{1.4cm}cP{1.4cm}cc@{}}
\toprule
Package                                                                      & Language  & Cross-sectional  & Temporal    & Cross-temporal  & Probabilistic & Multivariate \\
\midrule
\texttt{\href{https://pkg.earo.me/hts/}{hts}}
    & R         & \checkmark       &             &                 & \\
\texttt{\href{http://pkg.robjhyndman.com/thief/}{thief}}
    & R         &                  & \checkmark  &                 & \\
\texttt{\href{https://fable.tidyverts.org}{fable}}
    & R         & \checkmark       &             &                 & \checkmark\\
\texttt{\href{https://danigiro.github.io/FoReco/}{FoReco}}
    & R         & \checkmark       & \checkmark  & \checkmark      & \checkmark\\
\texttt{\href{https://cran.r-project.org/package=flap}{flap}}
    & R & & & & & \checkmark \\
\texttt{\href{https://angelpone.github.io/pyhts/}{pyhts}}
    & Python    & \checkmark       & \checkmark  &                 & \\
\texttt{\href{https://nixtla.github.io/hierarchicalforecast/}{hierarchicalforecast}}
    & Python    & \checkmark       &             &                 & \checkmark \\
\bottomrule
\end{tabular}

* `hts`, `thief`, and `FoReco` use `ts` objects
* `fable` uses `tsibble` objects
* `flap` uses matrices of base forecasts
* `fable` has plans to implement temporal and cross-temporal reconciliation

## Thanks!

\placefig{0}{1.2}{trim = 10 45 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{roman}
\placefig{2}{1.2}{trim = 0 40 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{george}
\placefig{4}{1.2}{trim = 0 10 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{hanlin}
\placefig{6}{1.2}{trim = 10 0 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{earowang}
\placefig{8}{1.2}{trim = 0 15 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{alanlee}
\placefig{10}{1.2}{trim = 30 0 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{mitch}
\placefig{12}{1.2}{trim = 15 0 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{shanika}
\placefig{14}{1.2}{trim = 0 0 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{tas}

\placefig{0}{3.8}{trim = 30 10 30 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{puwasala}
\placefig{2}{3.8}{trim = 0 10 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{fotios}
\placefig{4}{3.8}{trim = 0 30 30 20, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{nikos}
\placefig{6}{3.8}{trim = 50 30 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{souhaib}
\placefig{8}{3.8}{trim = 110 40 50 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{james}
\placefig{10}{3.8}{trim = 40 40 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{mahdi}
\placefig{12}{3.8}{trim = 50 50 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{christoph}
\placefig{14}{3.8}{trim = 50 50 0 20, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{fin}

\placefig{0}{6.4}{trim = 10 0 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{berwin}
\placefig{2}{6.4}{trim = 10 20 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{galit}
\placefig{4}{6.4}{trim = 10 0 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{mahsa}
\placefig{6}{6.4}{trim = 30 0 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{evan}
\placefig{8}{6.4}{trim = 5 25 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{bahman}
\placefig{10}{6.4}{trim = 0 40 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{pablo}
\placefig{12}{6.4}{trim = 0 40 0 0, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{danielegiro}
\placefig{14}{6.4}{trim = 0 0 0 30, clip=TRUE, width=10cm, height=2.5cm, keepaspectratio}{tommy}

## More information
\fontsize{18}{20}\sf

\href{https://robjhyndman.com}{\faicon{home} robjhyndman.com}

\href{https://aus.social/@robjhyndman}{\includegraphics[width=0.5cm]{figs/mastodon}\, aus.social/@robjhyndman}

\href{https://github.com/robjhyndman}{\faicon{github}  @robjhyndman}

\href{mailto:rob.hyndman@monash.edu}{\faicon{envelope}  rob.hyndman@monash.edu}

\vspace*{1cm}

\begin{block}{}
\centerline{\href{https://robjhyndman.com/curtin2024}{robjhyndman.com/curtin2024}}
\end{block}



\nocite{Di_FonGir2022a,temporal-hierarchies,ctprob}
\nocite{hierarchical,hfreview,coherentprob,htsgeometry,mint,flap}
